---
title: "MEP-LINCS Phase 2 GAL File Validation"
date: "`r Sys.Date()`"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,
                      echo=FALSE, warning=FALSE, message=FALSE)
```


```{r}
library("limma")#read GAL file and strsplit2
library("MEMA")#merge, annotate and normalize functions
library("data.table")#fast file reads, data merges and subsetting
library(ggplot2)
library(plotly)

source("MEPLINCSFunctions.R")

```


```{r processGALfile}
 # Read and clean spotmetadata
  
  #Read in the spot metadata from the gal file
  spotMetadata <- readSpotMetadata("GALFiles/20160120_LI8X001_2.gal")
  #Relabel the column Name to ECMpAnnotID
  setnames(spotMetadata, "Name", "ECMpAnnotID")
  
  #Add the print order and deposition number to the metadata
  #ldf <- readLogData("GALFiles/20160113.LI8X005.2.report.xml")
  #spotMetadata <- merge(smd,ldf, all=TRUE)
  setkey(spotMetadata,Spot)
  #Make a rotated version of the spot metadata to match the print orientation
  spotMetadata180 <- rotateMetadata(spotMetadata)
  
```

##MEMA Layout
All MEMAs in the experiment are in separate wells and have the same design of 48 ECM proteins spotted in 35 rows and 20 columns. The proteins are randomly assigned. The upper left and bottom right corners of each MEMA are image fiducials in the 488nm channel and there are four blank spots for checking orientation in all channels.  

<br>


```{r Content Layout,fig.width=8.5, fig.height=7}

#Select the A row wells and delete the blanks
DT <- spotMetadata
DT <- DT[!grepl("PBS",DT$ECMpAnnotID)]
p <- ggplot(DT,aes(x = ArrayColumn, y = ArrayRow, fill=ECMpAnnotID))+
  geom_point(shape=21, size = 3.5)+
  guides(fill=guide_legend(ncol = 2))+
  theme(legend.text = element_text(size = rel(.5)),legend.title=element_text(size = rel(1)),plot.title=element_text(size = rel(1)))+
  scale_y_reverse()+
  xlab("")+ylab("")+
  ggtitle(" \n\nLINCS MEMA A Row Layout")
print(p)

DT <- spotMetadata180
DT <- DT[!grepl("PBS",DT$ECMpAnnotID)]
p <- ggplot(DT,aes(x = ArrayColumn, y = ArrayRow, fill=ECMpAnnotID))+
  geom_point(shape=21, size = 3.5)+
  guides(fill=guide_legend(ncol = 2))+
  theme(legend.text = element_text(size = rel(.5)),legend.title=element_text(size = rel(1)),plot.title=element_text(size = rel(1)))+
  scale_y_reverse()+
  xlab("")+ylab("")+
  ggtitle(" \n\nLINCS MEMA B Row Layout")
print(p)
```


```{r Layout Replicate Count,echo=FALSE, message=FALSE, warnings=FALSE, fig.width=6.5, fig.height=3}
#Remove the fiducial and blank entries
setkey(DT,ECMpAnnotID)
DT <- DT[!"Fiducial"]
```


##Replicate Count
The average replicate count is `r  sprintf("%3.1f",mean(table(DT$ECMpAnnotID)))` and the range of the replicate counts is `r range(table(DT$ECMpAnnotID))`.  


```{r}

p <- ggplot(DT, aes(x=ECMpAnnotID))+
  geom_bar(width=.8)+geom_hline(yintercept = mean(table(DT$ECMpAnnotID)), colour="blue")+
  ggtitle(" \n\nCount of Replicate ECM Proteins In Each MEMA")+
  xlab("Printed ECM Protein")+ylab("Number of spots")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)),axis.title.x = element_text(size=rel(.8)),axis.title.y = element_text(size=rel(.8)),plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.5)))

print(p)

```
