---
output: html_document
---

---
title: "MEP-LINCS_PC3_Integrated_Analysis"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,
                      echo=FALSE, warning=FALSE, message=FALSE)
```


```{r }
#Author: Mark Dane, copyright 2015

source("MEPLINCSFunctions.R")
  
#Set the cell line to be analyzed (PC3|MCF7|YAPC)
cellLine <- "PC3"


```




```{r setup}
library("ggplot2")
library("data.table")
library("MEMA")
library("grid")
library(knitr)
library("gplots")
library("RColorBrewer")


#Setup colors for Barcode and text in all heatmaps
selDark2 <- colorRampPalette(brewer.pal(8,"Dark2"))
plateCol = selDark2(8)
ligandTypeCols = selDark2(28)
selReds <- colorRampPalette(brewer.pal(9,"Reds"))
redsPal <- selReds(10)

```


###Biological Feature Heat Maps
The following heat maps show PC3 cellline responses to MEPs across the staining sets. MEPs are across the rows, the image-derived features are in the columns. All values are normalized to the FBS control wells in their plates and median summarized across their ~15 replicates.  

Each MEP's set of normalized, summarized responses are curated to biologically relevent feature sets. These feature set vectors are clustered based on similarity.  


```{r, fig.height=16, eval=TRUE}
#Create a data.table with MEP rows, feature vector and metadata columns
#MEP rows should be median normalized
#Metadata in columns should be ECMp, Ligand, MEP
#Start with curated feature vectors that are unique
#Maybe summarize common parameters such as DNA content
#

#Start with the full combined dataset, no FBS
DT <- combineSSs(SSs = c("SS1","SS2","SS3"))

#Delete NID1
DT <- DT[!grepl("NID1",DT$ECMp)]
DT <- DT[!grepl("FBS",DT$Ligand)]

#Add a dummy values for ligand type
setkey(DT,Ligand)
DT$LigandType <- as.factor(toupper(substring(DT$Ligand, 1,2)))

#Build the feature set to keep:Normed values and MEP info
fullBio <- DT[,grep("RZSNorm|MEP|Ligand|ECMp",colnames(DT), value=TRUE), with = FALSE]
#Delete repeated columns, keeping the first one only
#TODO: Take medians of parameters that are common across staining sets
#such as SpotCellCount, Area, perimeter, DNA4n
#Reduce to biologically likely relevant
fullBio <- fullBio[,grep("[.](1|2)|Cells|Cytoplasm.*Dapi|Cytoplasm.*Area|CellMask|Center|Nuclei.*KRT|_SE|Euler", colnames(fullBio), value = TRUE, invert = TRUE), with = FALSE]
#Reduce to biologically minimal choice
fullBio <- fullBio[,grep("Actin|Integrated|Dapi", colnames(fullBio), value = TRUE, invert = TRUE), with = FALSE]

# Minimize column names
fullBio <- setnames(fullBio,colnames(fullBio),gsub("_RZSNorm|.*_CP_|.*_PA_|Intensity_|AreaShape_|Cycle_|Gated_","",colnames(fullBio)))

#Remove redundant features
reducedBio <- fullBio[,grep("Perimeter|4N|KRT|SpotCellCount|MedianEdu|H3", colnames(fullBio),value=TRUE, invert = TRUE),with=FALSE]

mreduced <- as.matrix(reducedBio[,grep("Ligand|ECMp|MEP",x = colnames(reducedBio), value = TRUE, invert = TRUE), with = FALSE])
rownames(mreduced) <- reducedBio$MEP

ms <- scale(mreduced)
#pdf("tmp")
full <- heatmap.2(ms, col = bluered, scale = "none", trace = "none", cexRow=.5, cexCol=1, cex.main=.5, key=TRUE, keysize = 1.5, key.title = NA, key.ylab = NA, density.info="none", 
          main = paste("PC3 Biological Features"), lhei=c(1.5,10.0), lwid=c(.2,1.5),mar=c(15,5))

```

```{r, fig.height=10, fig.width=14, eval=FALSE}

heatmap.2(ms, Rowv=full$rowDendrogram[[2]][[2]], breaks = full$breaks, col = bluered, trace = "none", cexRow=1, cexCol=1.5, cex.main=.5, key=TRUE, keysize = .1, key.title = NA, key.ylab = NA, density.info="none", 
          main = paste("PC3 Biological Features"), lhei=c(.8,5.0), lwid=c(.1,1.5),mar=c(14,9))  

#dev.off()
```

###Ligand and MEP Feature Vector Distance Heat Map

Method   
Combine all normalized and summarized intensity, morphology and spot values from all MEPS in the three staining sets  
Median summarize the control (FBS) responses on a per feature basis  
Compute the euclidean distance in feature space  between each MEP and the control  


```{r MEPDistanceHeatMap, fig.height=10, fig.width=14,eval=FALSE}

#Get a datatable of the normalized values across the staining set
SSs = c("SS1","SS2","SS3")
PC3DT <- integrateSSs(SSs, cellLine = "PC3")

#Calculate the distance of the numeric parameters from the FBS values

#Normalize the feature vectors to the medians
#to equally weight all features
setkey(PC3DT,"MEP")
parms <- colnames(PC3DT)[unlist(lapply(PC3DT,class)) %in% c("numeric","integer")]
parms <- grep("Center|_SE",parms,value = TRUE, invert = TRUE)
PC3DTNorm <- rbindlist(apply(PC3DT[,parms, with=FALSE], 1, scaleToMedians, normBase = PC3DT["FBS",parms,with=FALSE]))

#Calculate the dist matrix with euclidean method
dmm <- as.matrix(dist(PC3DTNorm), labels=TRUE)
#Extract the distance to the high serum medians
distToFBS <- dmm[which(PC3DT$MEP == "FBS"),]
#Name the distance values
names(distToFBS) <- PC3DT$MEP

distDT <- data.table(Dist = distToFBS, MEP =PC3DT$MEP, ECMp = PC3DT$ECMp, Ligand = PC3DT$Ligand)


#Need to remove FBS row to make into a rectangulat matrix
dFBS <- distToFBS[!names(distToFBS)=="FBS"]
#######This code needs logic that deals with missing ligands and ECMps

distM <- matrix(dFBS,nrow=length(unique(PC3DT$Ligand)))
setkey(PC3DT,ECMp)

colnames(distM) <-  unique(PC3DT$ECMp[!is.na(PC3DT$ECMp)])
rownames(distM) <- unique(PC3DT$LigandAnnotID[!PC3DT$Ligand == "FBS"])
#Plot full bio heatmap
#distBreaks <- seq(3.5,6,.1)
distBreaks <- seq(1500,12000,250)
heatmap.2(distM,  col = selReds, breaks = distBreaks, trace = "none", cexRow=.8, cexCol=.8, cex.main=1, key=TRUE, keysize = .1, key.title = NA, key.ylab = NA, density.info="none", 
          main = paste("MEP Distance From Control"), lhei=c(.8,5.0), lwid=c(.1,1.5),mar=c(14,9)) 


```

##EdU Response and Lineage Ratio

The following plots ranks each MEP along the x axis by their lineage ratio and EdU proliferation values. These two signals are combined in the last plot but it does not show a strong correlation.

```{r, eval=TRUE}

browser()
#Get a datatable of the normalized values across the staining set
SSs = c("SS1","SS2","SS3")
PC3DT <- integrateSSs(SSs, cellLine = "PC3")

#Remove low cell count spots
PC3DT <- PC3DT[PC3DT$Spot_PA_SpotCellCount_RZSNorm > -2]

p <- ggplot(PC3DT, aes(x =reorder(MEP, Cytoplasm_PA_Intensity_LineageRatio_RZSNorm), y = Cytoplasm_PA_Intensity_LineageRatio_RZSNorm))+
  geom_errorbar(aes(ymin=Cytoplasm_PA_Intensity_LineageRatio_RZSNorm-Cytoplasm_PA_Intensity_LineageRatio_RZSNorm_SE, ymax=Cytoplasm_PA_Intensity_LineageRatio_RZSNorm+Cytoplasm_PA_Intensity_LineageRatio_RZSNorm_SE), width=.001) +
xlab("MEP")+ylab("Lineage Ratio")+
   geom_point(colour = "blue",size = 3)+
  theme( axis.text.x=element_blank(), axis.ticks=element_blank(),  panel.grid.major = element_blank())+
   ggtitle("MEPs Ordered by Keratin Lineage Ratio with SE Bars")
print(p)

PC3DTLRExtremes <- PC3DT[PC3DT$Cytoplasm_PA_Intensity_LineageRatio_RZSNorm > quantile(PC3DT$Cytoplasm_PA_Intensity_LineageRatio_RZSNorm,.99) | PC3DT$Cytoplasm_PA_Intensity_LineageRatio_RZSNorm < quantile(PC3DT$Cytoplasm_PA_Intensity_LineageRatio_RZSNorm,.01)]

setkey(PC3DTLRExtremes,Cytoplasm_PA_Intensity_LineageRatio_RZSNorm)

kable(data.frame(PC3DTLRExtremes$MEP,PC3DTLRExtremes$Cytoplasm_PA_Intensity_LineageRatio_RZSNorm, PC3DTLRExtremes$Cytoplasm_PA_Intensity_LineageRatio_RZSNorm_SE), col.names = c("MEP","Keratin Lineage Ratio", "Standard Error"), digits = 2, align = c("l"))

p <- ggplot(PC3DT, aes(x =reorder(MEP, Nuclei_PA_Gated_EduPositiveProportion_RZSNorm), y = Nuclei_PA_Gated_EduPositiveProportion_RZSNorm))+
  geom_errorbar(aes(ymin=Nuclei_PA_Gated_EduPositiveProportion_RZSNorm-Nuclei_PA_Gated_EduPositiveProportion_RZSNorm_SE, ymax=Nuclei_PA_Gated_EduPositiveProportion_RZSNorm+Nuclei_PA_Gated_EduPositiveProportion_RZSNorm_SE), width=.01) +
xlab("MEP")+ylab("Normalized EdU Positive Proportion")+
    geom_point(colour = "red", size = 3)+
  theme( axis.text.x=element_blank(), axis.ticks=element_blank(),  panel.grid.major = element_blank())+
  ggtitle("MEPs Ordered by EdU Positive Proportion with SE Bars")
print(p)

PC3DTEPPExtremes <- PC3DT[PC3DT$Nuclei_PA_Gated_EduPositiveProportion_RZSNorm > quantile(PC3DT$Nuclei_PA_Gated_EduPositiveProportion_RZSNorm,.99) | PC3DT$Nuclei_PA_Gated_EduPositiveProportion_RZSNorm < quantile(PC3DT$Nuclei_PA_Gated_EduPositiveProportion_RZSNorm,.01)]

setkey(PC3DTEPPExtremes,Nuclei_PA_Gated_EduPositiveProportion_RZSNorm)

kable(data.frame(PC3DTEPPExtremes$MEP,PC3DTEPPExtremes$Nuclei_PA_Gated_EduPositiveProportion_RZSNorm,PC3DTEPPExtremes$Nuclei_PA_Gated_EduPositiveProportion_RZSNorm_SE), col.names = c("MEP","EdU Positive Proportion", "Standard Error"), digits = 2, align = c("l"))

p <- ggplot(PC3DT, aes(x =reorder(MEP, Nuclei_PA_Gated_EduPositiveProportion_RZSNorm), y = Cytoplasm_PA_Intensity_LineageRatio_RZSNorm))+geom_point()+
xlab("Rank Ordered EdU Positive Proportion")+ylab("Keratin Lineage Ratio")+
  theme( axis.text.x=element_blank(), axis.ticks=element_blank(),  panel.grid.major = element_blank())+
  ggtitle("Keratin Lineage Ratio by Ordered EdU Positive Proportion")
print(p)


p <- ggplot(PC3DT[PC3DT$Spot_PA_SpotCellCount_RZSNorm > -1.8], aes(x =Nuclei_PA_Gated_EduPositiveProportion_RZSNorm, y = Cytoplasm_PA_Intensity_LineageRatio_RZSNorm))+geom_point()+
xlab("Normalized EdU Positive Proportion")+ylab("Normalized Keratin Lineage Ratio")+
  theme( panel.grid.major = element_blank())+
  ggtitle("Keratin Lineage vs. EdU Positive Proportion\nFiltered for Spot Cell Count")

print(p)

```

```{r LowLineageLowEdUpp}
#Load level 2 data to get spot locations
l2 <- fread(paste0("./",cellLine,"/SS2/AnnotatedData/",cellLine,"_SS2_Level2.txt"), showProgress = FALSE)
l2$Spot_PA_OuterCell <- as.logical(gsub(" ","",l2$Spot_PA_OuterCell))
l2$Spot_PA_Sparse <- as.logical(gsub(" ","",l2$Spot_PA_Sparse))
l2$Spot_PA_Perimeter <- as.logical(gsub(" ","",l2$Spot_PA_Perimeter))

barcodes <- unique(l2$Barcode)
lthresh <- 0.6

DTl2 <- copy(l2)
#Remove the fiducial entries
setkey(DTl2,ECMp)
DTl2 <- DTl2[!"fiducial"]
DTl2 <- DTl2[!"blank"]
DTl2 <- DTl2[!"NID1"]

DT <- PC3DT[PC3DT$Spot_PA_SpotCellCount_RZSNorm > -2]

lowLineageLowEdUPPDT <- DT[(DT$Nuclei_PA_Gated_EduPositiveProportion_RZSNorm < -1.9) & (DT$Cytoplasm_PA_Intensity_LineageRatio_RZSNorm < -.375)]
kable(lowLineageLowEdUPPDT[,list(MEP)], caption = "MEPs with low EdU Positive Proportion and Low Lineage Ratio")


```


##Supplemental Material

###Spot Cell Count of FBS across Staining Sets

Each of the 56 ligands is paired with each of the 46 ECM proteins creating 2576 different MEPs. The plots below show the  MEPs median Spot Cell Counts for the 46 FBS MEPs in each staining set. Raw and normalized (median centered) values along with R squared values are shown.  

```{r,fig.height= 3.5, fig.width=3.5, eval=FALSE}


DT <- combineSSs(SSs = c("SS1","SS2","SS3"))[,list(MEP, Spot_PA_SpotCellCount,Spot_PA_SpotCellCount_RZSNorm, Spot_PA_SpotCellCount.1,Spot_PA_SpotCellCount_RZSNorm.1, Spot_PA_SpotCellCount.2,Spot_PA_SpotCellCount_RZSNorm.2)]

DT <- DT[grepl("FBS", DT$MEP)]

DT <- DT[,lapply(.SD, numericMedian), by="MEP"]

#Show scattter plots of raw and normalized spot cell count in each ss
p <- ggplot(DT, aes(x = Spot_PA_SpotCellCount, y = Spot_PA_SpotCellCount.1))+
  geom_point(siz = rel(.4), alpha = .6)+
  stat_smooth(method = "lm")+
  xlim(0,40)+ylim(0,40)+
  xlab("SS1 Spot Cell Count")+ylab("SS2 Spot Cell Count")+
  ggtitle("Spot Cell Count Across\nStaining Sets")+
  geom_text(x = 20, y = 35, label = lm_eqn(DT[,list(x=Spot_PA_SpotCellCount,  y=Spot_PA_SpotCellCount.1)]), parse = TRUE)
print(p)

p <- ggplot(DT, aes(x = Spot_PA_SpotCellCount_RZSNorm, y = Spot_PA_SpotCellCount_RZSNorm.1))+
  geom_point(siz = rel(.4), alpha = .6)+
  stat_smooth(method = "lm")+
  xlim(0,1.5)+ylim(0,1.5)+
  xlab("SS1 Spot Cell Count")+ylab("SS2 Spot Cell Count")+
  ggtitle("Normalized Spot Cell Count\nAcross Staining Sets")+
  geom_text(x = .7, y = 1.2, label = lm_eqn(DT[,list(x=Spot_PA_SpotCellCount_RZSNorm,  y=Spot_PA_SpotCellCount_RZSNorm.1)]), parse = TRUE)
print(p)

p <- ggplot(DT, aes(x = Spot_PA_SpotCellCount, y = Spot_PA_SpotCellCount.2))+ 
  geom_point(siz = rel(.4), alpha = .6)+
  stat_smooth(method = "lm")+
  xlim(0,40)+ylim(0,40)+
  xlab("SS1 Spot Cell Count")+ylab("SS3 Spot Cell Count")+
  ggtitle("Spot Cell Count \nAcross Staining Sets")+
  geom_text(x = 20, y = 35, label = lm_eqn(DT[,list(x=Spot_PA_SpotCellCount,  y=Spot_PA_SpotCellCount.2)]), parse = TRUE)
print(p)

p <- ggplot(DT, aes(x = Spot_PA_SpotCellCount_RZSNorm, y = Spot_PA_SpotCellCount_RZSNorm.2))+
  geom_point(siz = rel(.4), alpha = .6)+
stat_smooth(method = "lm")+
    xlim(0,1.5)+ylim(0,1.5)+
  xlab("SS1 Spot Cell Count")+ylab("SS3 Spot Cell Count")+
  ggtitle("Normalized Spot Cell Count \nAcross Staining Sets")+
  geom_text(x = .7, y = 1.2, label = lm_eqn(DT[,list(x=Spot_PA_SpotCellCount_RZSNorm,  y=Spot_PA_SpotCellCount_RZSNorm.2)]), parse = TRUE)
print(p)

p <- ggplot(DT, aes(x = Spot_PA_SpotCellCount.1, y = Spot_PA_SpotCellCount.2))+
  geom_point(siz = rel(.4), alpha = .6)+
stat_smooth(method = "lm")+
    xlim(0,40)+ylim(0,40)+
  xlab("SS2 Spot Cell Count")+ylab("SS3 Spot Cell Count")+
  ggtitle("Spot Cell Count \nAcross Staining Sets")+
  geom_text(x = 20, y = 35, label = lm_eqn(DT[,list(x=Spot_PA_SpotCellCount.1,  y=Spot_PA_SpotCellCount.2)]), parse = TRUE)
print(p)

p <- ggplot(DT, aes(x = Spot_PA_SpotCellCount_RZSNorm.1, y = Spot_PA_SpotCellCount_RZSNorm.2))+
  geom_point(siz = rel(.4), alpha = .6)+
stat_smooth(method = "lm")+
    xlim(0,1.5)+ylim(0,1.5)+
  xlab("SS2 Spot Cell Count")+ylab("SS3 Spot Cell Count")+
  ggtitle("Normalized Spot Cell Count \nAcross Staining Sets")+
  geom_text(x = .7, y = 1.2, label = lm_eqn(DT[,list(x=Spot_PA_SpotCellCount_RZSNorm.1,  y=Spot_PA_SpotCellCount_RZSNorm.2)]), parse = TRUE)
print(p)

```


###Spot Cell Count of FBS across Staining Sets

Each of the 56 ligands is paired with each of the 46 ECM proteins creating . The plots below show the MEPs median Spot Cell Counts for the all 56 ligands. Raw and normalized (median centered) values are shown. Coloring is by plate, each having seven ligand wells and one FBS control well.

```{r,fig.height= 3.5, fig.width=3.5, eval=FALSE}


DT <- combineSSs(SSs = c("SS1","SS2","SS3"))

#Show scattter plots of raw and normalized spot cell count in each ss
p <- ggplot(DT, aes(x = Spot_PA_SpotCellCount, y = Spot_PA_SpotCellCount.1, colour=Barcode))+
  geom_point(siz = rel(.4), alpha = .6)+
  stat_smooth(method = "loess",se=FALSE)+
  xlim(0,40)+ylim(0,40)+
  xlab("SS1 Spot Cell Count")+ylab("SS2 Spot Cell Count")+
  ggtitle("Spot Cell Count Across\nStaining Sets")+
  guides(colour=FALSE)
print(p)

p <- ggplot(DT, aes(x = Spot_PA_SpotCellCount_RZSNorm, y = Spot_PA_SpotCellCount_RZSNorm.1, colour=Barcode))+
  geom_point(siz = rel(.4), alpha = .6)+
  stat_smooth(method = "loess",se=FALSE)+
  xlim(0,1.5)+ylim(0,1.5)+
  xlab("SS1 Spot Cell Count")+ylab("SS2 Spot Cell Count")+
  ggtitle("Normalized Spot Cell Count\nAcross Staining Sets")+
  guides(colour=FALSE)
print(p)

p <- ggplot(DT, aes(x = Spot_PA_SpotCellCount, y = Spot_PA_SpotCellCount.2, colour=Barcode))+ 
  geom_point(siz = rel(.4), alpha = .6)+
  stat_smooth(method = "loess",se=FALSE)+
  xlim(0,40)+ylim(0,40)+
  xlab("SS1 Spot Cell Count")+ylab("SS3 Spot Cell Count")+
  ggtitle("Spot Cell Count \nAcross Staining Sets")+
  guides(colour=FALSE)
print(p)

p <- ggplot(DT, aes(x = Spot_PA_SpotCellCount_RZSNorm, y = Spot_PA_SpotCellCount_RZSNorm.2, colour=Barcode))+
  geom_point(siz = rel(.4), alpha = .6)+
  stat_smooth(method = "loess",se=FALSE)+
  xlim(0,1.5)+ylim(0,1.5)+
  xlab("SS1 Spot Cell Count")+ylab("SS3 Spot Cell Count")+
  ggtitle("Normalized Spot Cell Count \nAcross Staining Sets")+
  guides(colour=FALSE)
print(p)

p <- ggplot(DT, aes(x = Spot_PA_SpotCellCount.1, y = Spot_PA_SpotCellCount.2, colour=Barcode))+
  geom_point(siz = rel(.4), alpha = .6)+
  stat_smooth(method = "loess",se=FALSE)+
  xlim(0,40)+ylim(0,40)+
  xlab("SS2 Spot Cell Count")+ylab("SS3 Spot Cell Count")+
  ggtitle("Spot Cell Count \nAcross Staining Sets")+
  guides(colour=FALSE)
print(p)

p <- ggplot(DT, aes(x = Spot_PA_SpotCellCount_RZSNorm.1, y = Spot_PA_SpotCellCount_RZSNorm.2, colour=Barcode))+
  geom_point(siz = rel(.4), alpha = .6)+
  stat_smooth(method = "loess",se=FALSE)+
  xlim(0,1.5)+ylim(0,1.5)+
  xlab("SS2 Spot Cell Count")+ylab("SS3 Spot Cell Count")+
  ggtitle("Normalized Spot Cell Count \nAcross Staining Sets")+
  guides(colour=FALSE)
print(p)

```


```{r,fig.height= 3.5, fig.width=3.5, eval=FALSE}


DT <- combineSSs(SSs = c("SS1","SS2","SS3"))[,list(MEP, Spot_PA_SpotCellCount,Spot_PA_SpotCellCount_RZSNorm, Spot_PA_SpotCellCount.1,Spot_PA_SpotCellCount_RZSNorm.1, Spot_PA_SpotCellCount.2,Spot_PA_SpotCellCount_RZSNorm.2)]

DT <- DT[grepl("FBS", DT$MEP)]

DTS <- DT[,lapply(.SD, numericMedian), by="MEP"]

#Show scattter plots of raw and normalized spot cell count in each ss
p <- ggplot(DT, aes(x = Spot_PA_SpotCellCount, y = Spot_PA_SpotCellCount_RZSNorm, colour = MEP))+
  geom_point(siz = rel(.4), alpha = .6)+
  xlim(0,45)+ylim(0,1.3)+
  guides(colour = FALSE)+
  ggtitle("SS1 Spot Cell Count Normalization\nNot Summarized")
print(p)

#Show scattter plots of raw and normalized spot cell count in each ss
p <- ggplot(DTS, aes(x = Spot_PA_SpotCellCount, y = Spot_PA_SpotCellCount_RZSNorm))+
  geom_point(siz = rel(.4), alpha = .6)+
    stat_smooth(method = "lm")+
  xlim(0,45)+ylim(0,1.3)+
  ggtitle("SS1 Spot Cell CountNormalization\nSummarized")
print(p)

p <- ggplot(DT, aes(x = Spot_PA_SpotCellCount.1, y = Spot_PA_SpotCellCount_RZSNorm.1, colour = MEP))+
  geom_point(siz = rel(.4), alpha = .6)+
  xlim(0,45)+ylim(0,1.3)+
  guides(colour = FALSE)+
  ggtitle("SS2 Spot Cell Count Normalization\nNot Summarized")
print(p)

p <- ggplot(DTS, aes(x = Spot_PA_SpotCellCount.1, y = Spot_PA_SpotCellCount_RZSNorm.1))+
  geom_point(siz = rel(.4), alpha = .6)+
    stat_smooth(method = "lm")+
  xlim(0,45)+ylim(0,1.3)+
  ggtitle("SS2 Spot Cell Count Normalization\nSummarized")
print(p)

p <- ggplot(DT, aes(x = Spot_PA_SpotCellCount.2, y = Spot_PA_SpotCellCount_RZSNorm.2, colour = MEP))+
  geom_point(siz = rel(.4), alpha = .6)+
  xlim(0,45)+ylim(0,1.3)+
  guides(colour = FALSE)+ggtitle("SS3 Spot Cell Count Normalization\nNot Summarized")
print(p)

p <- ggplot(DTS, aes(x = Spot_PA_SpotCellCount.2, y = Spot_PA_SpotCellCount_RZSNorm.2))+ 
  geom_point(siz = rel(.4), alpha = .6)+
    stat_smooth(method = "lm")+
  xlim(0,45)+ylim(0,1.3)+
  ggtitle("SS3 Spot Cell Count Normalization\nSummarized")
print(p)


```