
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,
                      echo=FALSE, warning=FALSE, message=FALSE)
```

---
title: "MEP-LINCS Validation"
date: `r Sys.Date()`
output: html_document
---


```{r }
#Author: Mark Dane, copyright 2015

getl4HF <- function(cellLine, analysisVersion){
  ss="SS2"
  
  l3 <- fread(paste0("./",cellLine,"/",ss,"/AnnotatedData/",cellLine, "_",ss,"_",analysisVersion,"_Level3.txt"), showProgress = FALSE)
  
  #Set a threshold for filtering wells on their QA score
  wellQAThresh <- 0.7
  
  #TODO: Read this from Level 3 data
  lthresh <- 0.6
  
  #Number of PCS components to use
  nrPCs <- 9
  
  #Z score threshold for extreme spot cell count
  SCCZscoreThresh <- 3
  
  #Spot cell count threshold for HF dataset
  HFSCCThresh <- 20
  
  #Replicate count threshold for HF Dataset
  HFRepThresh <- 3
  
  #Remove the fiducial and blank data
  setkey(l3,ECMp)
  l3F <- l3[!"fiducial"]
  l3F <- l3F[!"blank"]
  setkey(l3F,"Ligand")
  l3F <- l3F[!grepl("FBS", l3F$Ligand)]
  
  #Remove failed QA wells
  l3F <- l3F[!l3F$QA_LowWellQA]
  #Filter for high spot occupancy and good quality DAPI####
  l3HF <- l3F[l3F$Spot_PA_SpotCellCount > HFSCCThresh]
  l3HF <- l3HF[!l3HF$QA_LowDAPIQuality]
  l3HF <- l3HF[!l3HF$QA_LowRegionCellCount]
  l4HF <- createl4(l3HF)
  l4HF <- l4HF[l4HF$Spot_PA_ReplicateCount >= HFRepThresh]
  return(l4HF)
}

```


```{r}
library(MEMA)
library(ggplot2)
library(plotly)
library(reshape2)
library(data.table)
library(DT)


validationMEPS <- function(cellLine, analysisVersion){
  ss="SS2"
#   
#   l3 <- fread(paste0("./",cellLine,"/",ss,"/AnnotatedData/",cellLine, "_",ss,"_",analysisVersion,"_Level3.txt"), showProgress = FALSE)
#   
#   barcodes <- sort(unique(l3$Barcode))
#   
#   
#   #Set a threshold for filtering wells on their QA score
#   wellQAThresh <- 0.7
#   
#   #TODO: Read this from Level 3 data
#   lthresh <- 0.6
#   
#   #Number of PCS components to use
#   nrPCs <- 9
#   
#   #Z score threshold for extreme spot cell count
#   SCCZscoreThresh <- 3
#   
#   #Spot cell count threshold for HF dataset
#   HFSCCThresh <- 20
#   
#   #Replicate count threshold for HF Dataset
#   HFRepThresh <- 3
#   
#   l3$OmeroDetailURL <- paste0('<a href="https://meplincs.ohsu.edu/webclient/img_detail/',l3$ImageID,'/"',' target="_blank">Omero</a>')
#   l3$OmeroThumbnailURL <- paste0('<a href="https://meplincs.ohsu.edu/webclient/render_thumbnail/',l3$ImageID,'/"',' target="_blank">Omero</a>')
#   l3$OmeroImageURL <- paste0('<a href="https://meplincs.ohsu.edu/webclient/render_image/',l3$ImageID,'/"',' target="_blank">Omero</a>')
#   
#   #Remove the fiducial and blank data
#   setkey(l3,ECMp)
#   l3F <- l3[!"fiducial"]
#   l3F <- l3F[!"blank"]
#   setkey(l3F,"Ligand")
#   l3F <- l3F[!grepl("FBS", l3F$Ligand)]
#   
#   #Remove failed QA wells
#   l3F <- l3F[!l3F$QA_LowWellQA]
#   #Filter for high spot occupancy and good quality DAPI####
#   l3HF <- l3F[l3F$Spot_PA_SpotCellCount > HFSCCThresh]
#   l3HF <- l3HF[!l3HF$QA_LowDAPIQuality]
#   l3HF <- l3HF[!l3HF$QA_LowRegionCellCount]
#   l4HF <- createl4(l3HF)
#   l4HF <- l4HF[l4HF$Spot_PA_ReplicateCount >= HFRepThresh]
  l4HF <- getl4HF(cellLine,analysisVersion)
  set.seed(1234)
  setkey(l4HF,Nuclei_PA_Gated_EduPositiveLogit_Norm)
  midMEPs <- round(quantile(1:nrow(l4HF), probs=c(.25,.75)))
  selectedMEPS <-l4HF[c(1:10, 
                        sample(seq(midMEPs[1], midMEPs[2], by=1), size = 5),
                        (nrow(l4HF)-9):nrow(l4HF))]
  
  return(selectedMEPS=selectedMEPS)
}
```


```{r, fig.width=6,fig.height=4.5}
PC3MEPS <- validationMEPS(cellLine = "PC3",analysisVersion = "v1.3")
PC3MEPS$MEP[PC3MEPS$MEP=="CDH8_KNG1"&PC3MEPS$CellLine=="PC3"] <- "NID1_FGF6"
PC3MEPS$ECMp[PC3MEPS$MEP=="NID1_FGF6"] <- "NID1"
PC3MEPS$Ligand[PC3MEPS$MEP=="NID1_FGF6"] <- "FGF6"
PC3MEPS$ECMp[PC3MEPS$MEP=="NID1_FGF6"] <- "NID1"
PC3MEPS$Nuclei_PA_Gated_EduPositiveLogit_Norm[PC3MEPS$MEP=="NID1_FGF6"] <- -7.17
PC3MEPS$SelectedFrom <- "PC3"

MCF7MEPS <- validationMEPS(cellLine = "MCF7",analysisVersion = "v1.3")
MCF7MEPS$SelectedFrom <- "MCF7"
MEPS <-setkey(rbind(PC3MEPS,MCF7MEPS),MEP)
```


```{r}
#write.csv(setkey(MEPS[,list(MEP,ECMp, Ligand, CellLine,Nuclei_PA_Gated_EduPositiveLogit_Norm,Barcode)],CellLine,Nuclei_PA_Gated_EduPositiveLogit_Norm), file = paste0("validationMEPS_v1.csv"))

#What are the EdU+ values for the 50 MEPs in the PC3 and MCF7 cell lines
#Merge the l4 HF MCF7 and PC3 datasets
l4HFDT <- rbindlist(lapply(c("PC3","MCF7"), getl4HF, analysisVersion="v1.3"))
#Plot the rank ordered EdU+ values, colored by cellline, shape by in set
l4HFDT$MC <-paste(l4HFDT$MEP,l4HFDT$CellLine,sep="_")
l4HFDT$Selected <- FALSE
l4HFDT$Selected[l4HFDT$MEP %in% MEPS$MEP] <- TRUE
setkey(l4HFDT,MEP)
setkey(MEPS,MEP)
l4HFDT <-merge(l4HFDT,MEPS[,list(MEP,SelectedFrom)],all=TRUE)
l4HFDT$ECMp[l4HFDT$MEP=="NID1_FGF6"] <- "NID1"
l4HFDT$Ligand[l4HFDT$MEP=="NID1_FGF6"] <- "FGF6"
l4HFDT$ECMp[l4HFDT$MEP=="NID1_FGF6"] <- "NID1"
l4HFDT$CellLine[l4HFDT$MEP=="NID1_FGF6"] <- "PC3"
l4HFDT$Selected[l4HFDT$MEP=="NID1_FGF6"] <- "TRUE"
l4HFDT$SelectedFrom[l4HFDT$MEP=="NID1_FGF6"] <- "PC3"
l4HFDT$Nuclei_PA_Gated_EduPositiveLogit_Norm[l4HFDT$MEP=="NID1_FGF6"] <- -7.17
```

###Method

The MEPs validation set is chosen from the highly filtered (HF) PC3 (`r paste0("n=",sum(l4HFDT$CellLine=="PC3"))`) and MCF7 (`r paste0("n=",sum(l4HFDT$CellLine=="MCF7"))`)  datasets. These datasets include responses from spots with 20 or more cells and 3 or more replicate spots. The validation set is selected from the RUV3 normalized EdU+ logit values.  


```{r}
#write.csv(setkey(l4HFDT[l4HFDT$Selected,list(MEP,ECMp, Ligand, CellLine,Nuclei_PA_Gated_EduPositiveLogit_Norm,Barcode)],CellLine,Nuclei_PA_Gated_EduPositiveLogit_Norm), file = paste0("validationMEPSInBothCellLines_v1.csv"))

for(cellLine in c("PC3","MCF7")){
  setkey(l4HFDT,CellLine)
  DT <- l4HFDT[cellLine]
  p <- ggplot(DT, aes(x =reorder(MEP, Nuclei_PA_Gated_EduPositiveLogit_Norm), y = Nuclei_PA_Gated_EduPositiveLogit_Norm, colour=SelectedFrom, shape=Selected, size=Selected))+
    xlab("MEP")+ylab("Normalized, EdU+ Proportion Ratio (logit)")+
    geom_point(alpha = 1)+
    theme( axis.text.x=element_blank(), axis.ticks=element_blank(),  panel.grid.major = element_blank())+
    ggtitle(paste("Validation MEPs in",cellLine,"Ordered by EdU+ Proportion"))
  
  print(p)
}
```

```{r MCF7vsPC3, fig.width=4.5,fig.height=4}
#Plot the responses in PC3 vs MCF7 for the MEPS validation set
setkey(l4HFDT,MEP)
setkey(MEPS,MEP)
MEPSDT <- l4HFDT[MEPS[,list(MEP)]]

MEPSDTc <- dcast(MEPSDT,MEP~CellLine,value.var = "Nuclei_PA_Gated_EduPositiveLogit_Norm")

p <- ggplot(MEPSDTc, aes(PC3,MCF7,colour=MEP))+
  geom_point()+
  coord_cartesian(xlim=c(-7,2),ylim=(c(-7,2)))+
  guides(colour=FALSE)+
  xlab("PC3 EdU+ (logit)")+
  ylab("MCF7 EdU+ (logit)")+
  ggtitle("EdU Responses in MCF7 vs PC3")

ggplotly(p)
```

###Validation MEP Table
The table below shows the normalized logits of the EdU positive proportions of all 50 MEPs in the MCF7 and PC3 cell lines.  

<br>

```{r datatable}
setkey(MEPSDT,Nuclei_PA_Gated_EduPositiveLogit_Norm)
datatable(format(MEPSDT[,list(MEP, CellLine, Nuclei_PA_Gated_EduPositiveLogit_Norm)], digits=2, scientific = FALSE, nsmall=0), options = list(pageLength = 5), colnames = c("MEP", "Cell Line","EDU+ Prop."))
```