
```{r "setup", include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE)

library(rmarkdown)
library(MEMA)
library(magrittr)
library(stringr)
library(parallel)
library(ggplot2)
library(RColorBrewer)

#Setup colors for Barcode and text in all heatmaps
selDark2 <- colorRampPalette(brewer.pal(8,"Dark2"))
plateCol = selDark2(8)
hmcols<-colorRampPalette(c("blue","white","red"))(16)
sampleProportion <- .5
```

```{r readLevel1Data}
barcodes <-getBarcodes(studyName, synId = "syn9946943")
l1DTL <- lapply(barcodes, function(barcode, path){
  sd <- fread(paste0(path,"/",barcode,"/Analysis/",barcode,"_Level1.tsv"),verbose = FALSE, showProgress = FALSE)
  set.seed(42)
  sd <- sd[sample(1:nrow(sd), size = .1*nrow(sd),replace = FALSE),]
}, path=path)
l1 <- rbindlist(l1DTL, fill=TRUE)
set.seed(42)
l1S <- l1[sample(1:nrow(l1), sampleProportion*nrow(l1), replace = FALSE),]

l1S <- l1S[,Well_Ligand :=  paste(Well,Ligand, sep="\n")]
if("Nuclei_PA_Cycle_State" %in% colnames(l1S)){
  l1S <- l1S[,TotalDNANormed := Nuclei_CP_Intensity_IntegratedIntensity_Dapi/median(Nuclei_CP_Intensity_IntegratedIntensity_Dapi[Nuclei_PA_Cycle_State==1]), by="Barcode"]
  l1S <- l1S[,DNAThresh := min(TotalDNANormed[Nuclei_PA_Cycle_State==2]), by="Barcode"]
}
#Determine number of rows in the plate for figures
nrPlateRow <- length(unique(str_extract(l1S$Well,"[[:alpha:]]")))
CellLine <- unique(l1S$CellLine)
```



#MEP-LINCS `r studyName` Cell-Level QA Report
####date: `r Sys.Date()`

##Summary
This report covers the cell-level data for the `r studyName` dataset. There are additional reports for the spot and MEP QA and the analysis of this staining set. 

##Introduction
The LINCS `r studyName` experiment was performed with `r CellLine` cells grown in `r length(barcodes)` 8-well plates. The  channels are `r  combine_words(gsub("Endpoint","",grep("Endpoint",colnames(l1S),value=TRUE)))`. The endpoints are `r combine_words(unique(l1S[,grep("Endpoint",colnames(l1S),value=TRUE), with=FALSE]))`. Color images of the cells at each spot were gathered on a Nikon automated microscope.

The histograms and scatterplots that display cell level data are from a randomly sampled subset of the full dataset.

<br>

###DAPI QA
Histograms of the total DAPI signal are typicaly bimodal with peaks at the 2N and 4N DNA positions. If the DAPI signal passed the QA test, a blue vertical line shows the gate value for classifying 2N and 4N populations.

For DAPI signals that have been gated, we want to normalize the DNA signal so that the G0/G1 peak has a value of 1. This should yield G2 peak at 2. To achieve this, the total DAPI intensity is divided by the median total DAPI intensity for all cells that have a DNA Cycle State value of 1 (to the left of the blue line).

The plots below are of a randomly sampled subset of the cell-level data.

<br>

```{r rawDAPIPlots, fig.width=3.7,fig.height=1.5*nrPlateRow}
cummulativeMaxCount<-0
if("TotalDNANormed" %in% colnames(l1S)){
  DAPISignal <- "TotalDNANormed"
  
  xaxisMax <- 4
} else {
  DAPISignal <- "Nuclei_CP_Intensity_IntegratedIntensity_Dapi"
  xaxisMax <- 100000000
}
for(barcode in barcodes){
  setkey(l1S, Barcode)
  dtb <- l1S[barcode]

  p <- ggplot(dtb, aes_string(x=DAPISignal))+geom_histogram(bins=300)+
    facet_grid(Barcode~Well_Ligand)
  currentMaxCount <- max(ggplot_build(p)$data[[1]]$count)
  cummulativeMaxCount<- max(cummulativeMaxCount,currentMaxCount)
}
for(barcode in barcodes){
  setkey(l1S, Barcode)
  dtb <- l1S[barcode]
  p <- ggplot(dtb, aes_string(x=DAPISignal))+geom_histogram(bins = 300)+
    coord_cartesian(x=c(0,xaxisMax), y=c(0,cummulativeMaxCount))+
    facet_wrap(~Well_Ligand, nrow=nrPlateRow) +
    ylab("Count")+xlab("Total Intensity DAPI")+ggtitle(paste("Total DAPI Intensity:",barcode))+
    theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5, size=rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.8)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)),strip.text.x = element_text(size = rel(.4)))
  if("TotalDNANormed" %in% colnames(l1S)) p <- p +  geom_vline(data = dtb, aes(xintercept = DNAThresh), colour = "blue")
  suppressWarnings(print(p))
}
```

<br>


```{r, child='MEP-LINCS_MitoText.Rmd', eval="Cytoplasm_CP_Intensity_MedianIntensity_MitoTracker" %in% colnames(l1S)}
```

<br>

```{r rawMitoHistograms, fig.width=3.7,fig.height=4, eval="Cytoplasm_CP_Intensity_MedianIntensity_MitoTracker" %in% colnames(l1S)}

for (barcode in barcodes){
  setkey(l1S, Barcode)
  dtb <- l1S[barcode]
  dtb <- dtb[log2(dtb$Cytoplasm_CP_Intensity_MedianIntensity_MitoTracker) > quantile(log2(dtb$Cytoplasm_CP_Intensity_MedianIntensity_MitoTracker), probs=.01, na.rm=TRUE) & log2(dtb$Cytoplasm_CP_Intensity_MedianIntensity_MitoTracker) < quantile(log2(dtb$Cytoplasm_CP_Intensity_MedianIntensity_MitoTracker),probs=.98, na.rm=TRUE)]
  p <- ggplot(dtb, aes(x=log2(Cytoplasm_CP_Intensity_MedianIntensity_MitoTracker)))+
    geom_histogram(bins = 300)+
    facet_wrap(~Well_Ligand, nrow=2, scales="free_x")+
    ggtitle(paste("MitoTracker Intensity,",barcode))+
    ylab("Count")+xlab("Median Intensity MitoTracker\nLog2")+
    theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust=0.5, size=rel(1)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.7)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)), strip.text = element_text(size = rel(.5)))
  suppressWarnings(print(p))
}
```

```{r rawCaspaseHistograms, fig.width=3.7,fig.height=4, eval="Cytoplasm_CP_Intensity_MedianIntensity_Caspase" %in% colnames(l1S)}

for (barcode in barcodes){
  setkey(l1S, Barcode)
  dtb <- l1S[barcode]
  dtb <- dtb[dtb$Cytoplasm_CP_Intensity_MedianIntensity_CaspaseLog2 > quantile(dtb$Cytoplasm_CP_Intensity_MedianIntensity_CaspaseLog2, probs=.01, na.rm=TRUE) & dtb$Cytoplasm_CP_Intensity_MedianIntensity_CaspaseLog2 < quantile(dtb$Cytoplasm_CP_Intensity_MedianIntensity_CaspaseLog2,probs=.98, na.rm=TRUE)]
  p <- ggplot(dtb, aes(x=log2(Cytoplasm_CP_Intensity_MedianIntensity_CaspaseLog2)))+
    geom_histogram(bins = 300)+
    facet_wrap(~Well_Ligand, nrow=2, scales="free_x")+
    ggtitle(paste("Caspase Intensity,",barcode))+
    ylab("Count")+xlab("Median Intensity Caspase\nLog2")+
    theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust=0.5, size=rel(1)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.7)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)), strip.text = element_text(size = rel(.5)))
  suppressWarnings(print(p))
}
```


```{r, child='MEP-LINCS_EdUText.Rmd', eval="Nuclei_CP_Intensity_MedianIntensity_EdU" %in% colnames(l1S)}
```

<br>

```{r rawEdUHistograms, fig.width=3.7,fig.height=4, eval="Nuclei_CP_Intensity_MedianIntensity_EdU" %in% colnames(l1S)}

for (barcode in barcodes){
  setkey(l1S,Barcode)
  mDT <- l1S[barcode]
  #Squish the extreme values for better display
  mDT <- mDT[mDT$Nuclei_CP_Intensity_MedianIntensity_EdU > quantile(mDT$Nuclei_CP_Intensity_MedianIntensity_EdU, probs=.01, na.rm=TRUE) & mDT$Nuclei_CP_Intensity_MedianIntensity_EdU < quantile(mDT$Nuclei_CP_Intensity_MedianIntensity_EdU,probs=.98, na.rm=TRUE),]
  p <- ggplot(mDT, aes(x=log2(Nuclei_CP_Intensity_MedianIntensity_EdU)))+
    geom_histogram(bins = 300)+
    facet_wrap(~Well_Ligand, nrow=2, scales="free_x")+
    ggtitle(paste("EdU Intensity,",barcode))+
    ylab("Count")+xlab("Median Intensity EdU\nLog2")+
    theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust=0.5, size=rel(1)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)), strip.text = element_text(size = rel(.5)))
  suppressWarnings(print(p))
}

```



```{r, child='MEP-LINCS_Krt5Krt19Text.Rmd', eval="Cytoplasm_CP_Intensity_MedianIntensity_KRT5" %in% colnames(l1S)}
```

<br>

```{r rawKrt5Histograms, fig.width=3.7,fig.height=4, eval="Cytoplasm_CP_Intensity_MedianIntensity_KRT5" %in% colnames(l1S)}
for (barcode in barcodes){
  mDT <- l1S[l1S$Barcode == barcode]
    #Squish the extreme values for better display
  mDT <- mDT[mDT$Cytoplasm_CP_Intensity_MedianIntensity_KRT5 > quantile(mDT$Cytoplasm_CP_Intensity_MedianIntensity_KRT5, probs=.01, na.rm=TRUE) & mDT$Cytoplasm_CP_Intensity_MedianIntensity_KRT5 < quantile(mDT$Cytoplasm_CP_Intensity_MedianIntensity_KRT5,probs=.98, na.rm=TRUE),]
  p <- ggplot(mDT, aes(x=log2(Cytoplasm_CP_Intensity_MedianIntensity_KRT5)))+
    geom_histogram(bins = 300)+
    facet_wrap(~Well_Ligand, nrow=2)+
    ggtitle(paste("KRT5 Intensity,",barcode))+
    ylab("Count")+xlab("Median Intensity KRT5\nLog2")+
    theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust=0.5, size=rel(1)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)), strip.text = element_text(size = rel(.5)))
  suppressWarnings(print(p))
}

```


<br>

```{r rawKrt19Histograms, fig.width=3.7,fig.height=4, eval="Cytoplasm_CP_Intensity_MedianIntensity_KRT19" %in% colnames(l1S)}
for (barcode in barcodes){
  mDT <- l1S[l1S$Barcode == barcode]
  #Squish the extreme values for better display
  mDT <- mDT[mDT$Cytoplasm_CP_Intensity_MedianIntensity_KRT19 > quantile(mDT$Cytoplasm_CP_Intensity_MedianIntensity_KRT19, probs=.01, na.rm=TRUE) & mDT$Cytoplasm_CP_Intensity_MedianIntensity_KRT19 < quantile(mDT$Cytoplasm_CP_Intensity_MedianIntensity_KRT19,probs=.98, na.rm=TRUE),]
  p <- ggplot(mDT, aes(x=log2(Cytoplasm_CP_Intensity_MedianIntensity_KRT19)))+
    geom_histogram(bins = 300)+
    facet_wrap(~Well_Ligand, nrow=2)+
    ggtitle(paste("KRT19 Intensity,",barcode))+
    ylab("Count")+xlab("Median Intensity KRT19\nLog2")+
    theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust=0.5, size=rel(1)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)), strip.text = element_text(size = rel(.5)))
  suppressWarnings(print(p))
}

```


<br>

```{r rawlineageRatioHistograms, fig.width=3.7,fig.height=4, eval="Cytoplasm_PA_Intensity_LineageRatio" %in% colnames(l1S)}
for (barcode in barcodes){
  mDT <- l1S[l1S$Barcode == barcode]
   mDT <- mDT[mDT$Cytoplasm_PA_Intensity_LineageRatio > quantile(mDT$Cytoplasm_PA_Intensity_LineageRatio, probs=.01, na.rm=TRUE) & mDT$Cytoplasm_PA_Intensity_LineageRatio < quantile(mDT$Cytoplasm_PA_Intensity_LineageRatio,probs=.98, na.rm=TRUE),]
  p <- ggplot(mDT, aes(x=log2(Cytoplasm_PA_Intensity_LineageRatio)))+
    geom_histogram(bins = 300)+
    facet_wrap(~Well_Ligand, nrow=2)+
    ggtitle(paste("Lineage Ratio,",barcode))+
    ylab("Count")+xlab("Median Intensity Lineage Ratio\nLog2")+
    theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5, size=rel(1)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.7)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)), strip.text = element_text(size = rel(.5)))
  suppressWarnings(print(p))
}

```


```{r gatedKRTScatterplots, fig.width=8,fig.height=4, eval=length(setdiff(c("Cytoplasm_CP_Intensity_MedianIntensity_KRT5","Cytoplasm_CP_Intensity_MedianIntensity_KRT19","Cytoplasm_PA_Gated_KRTClass"),colnames(l1S)))==0}

for (barcode in barcodes){
  mDT <- l1S[l1S$Barcode == barcode]
  p <- ggplot(mDT, aes(x=log2(Cytoplasm_CP_Intensity_MedianIntensity_KRT5),y=log2(Cytoplasm_CP_Intensity_MedianIntensity_KRT19), colour=factor(Cytoplasm_PA_Gated_KRTClass)))+
    geom_point(size=rel(.2), alpha=.2)+
    geom_rug(col=rgb(.5,0,0,alpha=.01))+
    facet_wrap(~Well_Ligand, nrow=2)+
    ggtitle(paste("KRT19 vs KRT5 Intensities,",barcode))+
    guides(colour=guide_legend(title="KRT Class", override.aes = list(alpha = 1, size=1)))+
    ylab("Median KRT19 Intensity\nLog2")+xlab("Median KRT5 Intensity\nLog2")+
    theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5, size=rel(.5)),axis.title.x = element_text(size = rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(.5)), axis.title.y = element_text(size = rel(.5)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.5)),legend.title=element_text(size = rel(.5)), strip.text = element_text(size = rel(.5)))
  suppressWarnings(print(p))
}

```


```{r EdUgatedKRTScatterplots, fig.width=8,fig.height=4, eval=length(setdiff(c("Cytoplasm_CP_Intensity_MedianIntensity_KRT19","Nuclei_CP_Intensity_MedianIntensity_EdU","Cells_PA_Gated_EdUKRT19Class"),colnames(l1S)))==0}

for (barcode in barcodes){
  mDT <- l1S[l1S$Barcode == barcode]
  p <- ggplot(mDT, aes(x=log2(Cytoplasm_CP_Intensity_MedianIntensity_KRT19),y=log2(Nuclei_CP_Intensity_MedianIntensity_EdU), colour=factor(Cells_PA_Gated_EdUKRT19Class)))+
    geom_point(size=rel(.2), alpha=.2)+
    geom_rug(col=rgb(.5,0,0,alpha=.01))+
    facet_wrap(~Well_Ligand, nrow=2)+
    ggtitle(paste("EdU vs KRT19 Intensities,",barcode))+
    guides(colour=guide_legend(title="KRT EdU Class", override.aes = list(alpha = 1, size=1)))+
    ylab("Median EdU Intensity\nLog2")+xlab("Median KRT19 Intensity\nLog2")+
    theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5, size=rel(.5)),axis.title.x = element_text(size = rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(.5)), axis.title.y = element_text(size = rel(.5)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.5)),legend.title=element_text(size = rel(.5)), strip.text = element_text(size = rel(.5)))
  suppressWarnings(print(p))
}

```

```{r EdUgatedKRT5Scatterplots, fig.width=8,fig.height=4, eval=length(setdiff(c("Cytoplasm_CP_Intensity_MedianIntensity_KRT5","Nuclei_CP_Intensity_MedianIntensity_EdU","Cells_PA_Gated_EdUKRT5Class"),colnames(l1S)))==0}

for (barcode in barcodes){
  mDT <- l1S[l1S$Barcode == barcode]
  p <- ggplot(mDT, aes(x=log2(Cytoplasm_CP_Intensity_MedianIntensity_KRT5),y=log2(Nuclei_CP_Intensity_MedianIntensity_EdU), colour=factor(Cells_PA_Gated_EdUKRT5Class)))+
    geom_point(size=rel(.2), alpha=.2)+
    geom_rug(col=rgb(.5,0,0,alpha=.01))+
    facet_wrap(~Well_Ligand, nrow=2)+
    ggtitle(paste("EdU vs KRT5 Intensities,",barcode))+
    guides(colour=guide_legend(title="KRT EdU Class", override.aes = list(alpha = 1, size=1)))+
    ylab("Median EdU Intensity\nLog2")+xlab("Median KRT5 Intensity\nLog2")+
    theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5, size=rel(.5)),axis.title.x = element_text(size = rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(.5)), axis.title.y = element_text(size = rel(.5)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.5)),legend.title=element_text(size = rel(.5)), strip.text = element_text(size = rel(.5)))
  suppressWarnings(print(p))
}

```

```{r EdUbygatedKRTLigandboxplots, fig.width=10,fig.height=4, eval=length(setdiff(c("Nuclei_CP_Intensity_MedianIntensity_EdU","Cytoplasm_PA_Gated_KRT19Positive"),colnames(l1S)))==0}

p <- ggplot(l1S, aes(x=factor(Ligand),y=log2(Nuclei_CP_Intensity_MedianIntensity_EdU), colour=factor(Cytoplasm_PA_Gated_KRT19Positive)))+
  geom_boxplot(outlier.colour = NA)+
  coord_cartesian(ylim = c(6,8))+
  ggtitle(paste("EdU Intensities by KRT19 State and Ligand"))+
  guides(colour=guide_legend(title="KRT19+", override.aes = list(alpha = 1, size=1)))+
  ylab("Median EdU Intensity\nLog2")+xlab("Ligand")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust=0.5, size=rel(.5)),axis.title.x = element_text(size = rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(.5)), axis.title.y = element_text(size = rel(.5)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.5)),legend.title=element_text(size = rel(.5)), strip.text = element_text(size = rel(.5)))
suppressWarnings(print(p))
```

```{r EdUbygatedKRT5Ligandboxplots, fig.width=10,fig.height=4, eval=length(setdiff(c("Nuclei_CP_Intensity_MedianIntensity_EdU","Cytoplasm_PA_Gated_KRT5Positive"),colnames(l1S)))==0}

p <- ggplot(l1S, aes(x=factor(Ligand),y=log2(Nuclei_CP_Intensity_MedianIntensity_EdU), colour=factor(Cytoplasm_PA_Gated_KRT5Positive)))+
  geom_boxplot(outlier.colour = NA)+
  coord_cartesian(ylim = c(6,8))+
  ggtitle(paste("EdU Intensities by KRT5 State and Ligand"))+
  guides(colour=guide_legend(title="KRT5+", override.aes = list(alpha = 1, size=1)))+
  ylab("Median EdU Intensity\nLog2")+xlab("Ligand")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust=0.5, size=rel(.5)),axis.title.x = element_text(size = rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(.5)), axis.title.y = element_text(size = rel(.5)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.5)),legend.title=element_text(size = rel(.5)), strip.text = element_text(size = rel(.5)))
suppressWarnings(print(p))
```

<br>

#Supplemental Material

```{r Cell_cycle_plots,echo=FALSE, fig.width=8, fig.height=5, eval=length(setdiff(c("TotalDNANormed","Nuclei_CP_Intensity_MedianIntensity_EdU","Nuclei_PA_Gated_EdUPositive"),colnames(l1S)))==0}

for (barcode in barcodes){
  setkey(l1S, Barcode)
  dtb <- l1S[barcode]
  p <- ggplot(dtb, aes(x=TotalDNANormed, y = log2(Nuclei_CP_Intensity_MedianIntensity_EdU), colour = Nuclei_PA_Gated_EdUPositive))+geom_point(size = rel(.2), alpha = .4)+
    geom_rug(col=rgb(.5,0,0,alpha=.01))+
    coord_cartesian(x=c(0,6))+
    scale_x_continuous(name="Total Intensity DAPI",breaks=0:6)+
    facet_wrap(~Well_Ligand, nrow=2)+ggtitle(paste("EdU vs. DAPI,",barcode))+ylab("EdU\n(log2)")+
    guides(colour=FALSE)+
    theme(strip.text = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
  suppressWarnings(print(p))
}
```

```{r KRT5EdU_plots,echo=FALSE, fig.width=8, fig.height=5, eval=length(setdiff(c("Cytoplasm_CP_Intensity_MedianIntensity_KRT5","Nuclei_CP_Intensity_MedianIntensity_EdU","Nuclei_PA_Gated_EdUPositive"),colnames(l1S)))==0}

for (barcode in barcodes){
  setkey(l1S, Barcode)
  dtb <- l1S[barcode]
  p <- ggplot(dtb, aes(x=log2(Cytoplasm_CP_Intensity_MedianIntensity_KRT5), y = log2(Nuclei_CP_Intensity_MedianIntensity_EdU), colour = Nuclei_PA_Gated_EdUPositive))+geom_point(size = rel(.2), alpha = .4)+
    geom_rug(col=rgb(.5,0,0,alpha=.01))+
    scale_x_continuous(name="Median Intensity KRT5 \n(log2)",breaks=0:6)+
    facet_wrap(~Well_Ligand, nrow=2)+ggtitle(paste("EdU vs. KRT5,",barcode))+ylab("EdU\n(log2)")+
    guides(colour=FALSE)+
    theme(strip.text = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
  suppressWarnings(print(p))
}
```

