
```{r "setup", include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,
                      echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE)

#Author: Mark Dane, copyright 2015-2017

library(MEMA)
library(magrittr)
library(parallel)
library(ggplot2)
# library("grid")
# library("knitr")
# library("gplots")
 library("RColorBrewer")
# library(DT)
# library(d3heatmap)
# library(plotly)

#Setup colors for Barcode and text in all heatmaps
selDark2 <- colorRampPalette(brewer.pal(8,"Dark2"))
plateCol = selDark2(8)
hmcols<-colorRampPalette(c("blue","white","red"))(16)
```


```{r readLevel3and4Data, echo=FALSE}

l3 <- fread(paste0(path,"/",studyName,"/Annotated/",studyName,"_Level3.tsv"),verbose = FALSE, showProgress = FALSE)
l4 <- fread(paste0(path,"/",studyName,"/Annotated/",studyName,"_Level4.tsv"),verbose = FALSE, showProgress = FALSE)

k <- unique(l3$k)
wellQAThresh <- unique(l3$QA_lowWellQAThreshold)
lowRegionCellCountThreshold <- unique(l3$QA_lowRegionCellCountThreshold)
lowSpotCellCountThreshold <-unique(l3$QA_lowSpotCellCountThreshold)
annotations <- fread(paste0(path,"/",studyName,"/Annotated/",studyName,"_Level3Annotations.tsv"),header = FALSE)
CellLine = annotations$V2[annotations$V1=="CellLine"]
Preprocess = annotations$V2[annotations$V1=="Preprocess"]
DataType = annotations$V2[annotations$V1=="DataType"]
Consortia = annotations$V2[annotations$V1=="Consortia"]
Drug = annotations$V2[annotations$V1=="Drug"]
Segmentation = annotations$V2[annotations$V1=="Segmentation"]
StainingSet = annotations$V2[annotations$V1=="StainingSet"]
Level = annotations$V2[annotations$V1=="Level"]

#Use imageIDs to add URLs
if(!Segmentation=="v1"){
  l3$OmeroDetailURL <- paste0('<a href="https://meplincs.ohsu.edu/webclient/img_detail/',l3$ImageID,'/"',' target="_blank">Omero</a>')
  l3$OmeroThumbnailURL <- paste0('<a href="https://meplincs.ohsu.edu/webclient/render_thumbnail/',l3$ImageID,'/"',' target="_blank">Omero</a>')
  l3$OmeroImageURL <- paste0('<a href="https://meplincs.ohsu.edu/webclient/render_image/',l3$ImageID,'/"',' target="_blank">Omero</a>')
}
barcodes <- unique(l3$Barcode)
l3 <- l3[,Well_Ligand :=  paste(Well,Ligand, sep="\n")]

```


##MEP-LINCS Spot Level Quality Assessment Report
## `r studyName` 
####date: `r Sys.Date()`

###Summary
This report provides Quality Assessment (QA) of the `r studyName` MEP-LINCS Microenvironment Microarray (MEMA) data. After QA filtering, there are `r nrow(l4[!grepl("PBS|Fiducial",l4$ECMp),][!grepl("FBS",l4$Ligand)])` Microenvironment Perturbations (MEPs) that are pairwise combinations of `r length(unique(l4$ECMp[!grepl("PBS|Fiducial",l4$ECMp)]))` ECM proteins and `r length(unique(l4$Ligand[!grepl("FBS",l4$Ligand)]))` ligands or growth factors.  

###Introduction
The MEP-LINCS `r CellLine` `r StainingSet` experiment was performed with cells grown in `r length(barcodes)` 8-well plates. The `r StainingSet` staining set channels are `r  combine_words(gsub("Endpoint","",grep("Endpoint",colnames(l3),value=TRUE)))`. The endpoints are `r combine_words(unique(l3[,grep("Endpoint",colnames(l3),value=TRUE), with=FALSE]))`. Color images of the cells at each spot were gathered on a Nikon automated microscope. The Cell Profiler Pipeline version `r Segmentation` performed the segmentation and the analysis pipeline version is `r Preprocess`.

Intensity, position and morphology data are gathered for the nucleus and cytoplasm of each cell, merged with the experiment metadata, normalized, filtered and summarized. The dataset is organized into the four LINCS imaging categories as follows:

Level 1 - Raw, annotated cell level data  
Level 2 - Normalized, annotated cell data  
Level 3 - Normalized, annotated data summarized to the spot level  
Level 4 - Normalized, annotated data summarized to the replicate (MEP) level  

The data merging and analysis is done in R using open source software and is available at the GitHub repository https://github.com/MEP-LINCS/MEP_LINCS. 

<br>

###QA Scoring of the dataset
Each well is scored for even cell seeding according to the count of the DAPI-stained nuclei. A detailed explanation of the QA method is in the supplemental material. In brief, cell counts at the spot level and locally-averaged neighborhoods are used to score the wells and filter the dataset. QA Scores range from 0 to 1 and represent the proportion of the spots that have at least one cell and are not in low cell count neighborhoods.

The following plots are pseudo images of each MEMA's spot cell count along with histograms of the loess model with its QA score. 


```{r Heatmaps_QAScores, fig.width=3.7,fig.height=4}
plotSCCHeatmapsQAHistograms(l3, barcodes, lthresh=wellQAThresh)

```

The following histograms show the raw spot cell counts for each array.  

```{r SCCHistograms, fig.width=3.7,fig.height=4, eval=TRUE}
xLimits <- quantile(l3$Spot_PA_SpotCellCount, probs=c(0,.995))
yLimits <- c(0,75)

for (barcode in barcodes){
  setkey(l3,"Barcode")
  dt <- l3[barcode]
  dt$Well_Ligand <- sub("_","\n",dt$Well_Ligand)
  
  p <- ggplot(dt, aes(x=Spot_PA_SpotCellCount))+
    geom_histogram(binwidth = 1)+
    coord_cartesian(xlim=xLimits, ylim=yLimits)+
    ggtitle(paste("\n Spot Cell Count Histogram of plate",unique(dt$Barcode)))+
    xlab("Spot Cell Count")+
    theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(.5)),
          axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(.7)),
          plot.title = element_text(size = rel(.8)),
          strip.text = element_text(size = rel(.5)),
          legend.text=element_text(size = rel(.3)),
          legend.title=element_text(size = rel(.3)))+
    facet_wrap(~Well_Ligand, ncol=4)
  suppressWarnings(print(p))
  
}

```

<br>

###Spot Cell Count Boxplots
Boxplots of the spot cell counts show the values stratified by plate and well. The blue line shows the dataset median. Each boxplot summarizes data from ~700 spots.     

```{r SCCBoxplots, fig.height=5}

p <- ggplot(l3, aes(x=factor(Barcode), y=Spot_PA_SpotCellCount, colour=Well))+
  geom_boxplot(outlier.size = .5)+
  xlab("")+ylab("Spot Cell Count")+
  ggtitle("Spot Cell Count by Well and Barcode ")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust=0.5, size=rel(1)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.8)))
p <- p + geom_hline(yintercept = median(l3$Spot_PA_SpotCellCount), colour="blue", alpha=.5)
print(p)

```

```{r SCCBoxplotsbyLigand, fig.height=5}

p <- ggplot(l3, aes(x=factor(Ligand), y=Spot_PA_SpotCellCount, colour=factor(Barcode)))+
  geom_boxplot(outlier.size = .5)+
  labs(x="Ligand",y="Spot Cell Count",colour="Barcode", title="Spot Cell Count by Ligand and Barcode")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust=0.5, size=rel(1)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.8)))
p <- p + geom_hline(yintercept = median(l3$Spot_PA_SpotCellCount), colour="blue", alpha=.5)
print(p)

```

###DAPI Boxplots

DAPI stains the DNA of each nucleus and is used to idenitify the nuclear regions and quantify the amount of DNA. Boxplots of the DAPI intensities show the values stratified by plate and well. Each boxplot summarizes data from ~700 spots.  

```{r DAPIBoxplots, fig.height=5}

p <- ggplot(l3, aes(x=factor(Barcode), y=log2(Nuclei_CP_Intensity_MedianIntensity_Dapi), colour=Well))+
  geom_boxplot(outlier.size = .5)+
  labs(x="Ligand",y="Median Intensity DAPI (log2)",colour="Well", title="Median DAPI Intensity per Spot by Well and Barcode")+
  theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5, size=rel(1)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.8)))
p <- p + geom_hline(yintercept = log2(median(l3$Nuclei_CP_Intensity_MedianIntensity_Dapi)), colour="blue", alpha=.5)
print(p)

```


```{r, child='MEP-LINCS_MitoText.Rmd', eval="Cytoplasm_CP_Intensity_MedianIntensity_MitoTracker" %in% colnames(l3)}
```



```{r rawMitoBoxplots, fig.height=5,  eval="Cytoplasm_CP_Intensity_MedianIntensity_MitoTracker" %in% colnames(l3)}

p <- ggplot(l3, aes(x=factor(Barcode), y=log2(Cytoplasm_CP_Intensity_MedianIntensity_MitoTracker), colour=Well))+
  geom_boxplot(outlier.size = .5)+
  xlab("")+ylab("Median Intensity MitoTracker (log2)")+
  ggtitle("Median Intensity MitoTracker per Spot by Well and Barcode ")+
  theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5, size=rel(1)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.8)))
p <- p + geom_hline(yintercept = log2(median(l3$Cytoplasm_CP_Intensity_MedianIntensity_MitoTracker)), colour="blue", alpha=.5)
print(p)

```



```{r rawcaspaseBoxplots, fig.height=5,  eval="Cytoplasm_CP_Intensity_MedianIntensity_Caspase" %in% colnames(l3)}

p <- ggplot(l3, aes(x=factor(Barcode), y=log2(Cytoplasm_CP_Intensity_MedianIntensity_Caspase), colour=Well))+
  geom_boxplot(outlier.size = .5)+
  xlab("")+ylab("Median Intensity Caspase (log2)")+
  ggtitle("Median Intensity Caspase per Spot by Well and Barcode ")+
  theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5, size=rel(1)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.8)))
p <- p + geom_hline(yintercept = median(l3$Cytoplasm_CP_Intensity_MedianIntensity_CaspaseLog2), colour="blue", alpha=.5)
print(p)

```


```{r, child='MEP-LINCS_EdUText.Rmd', eval="Nuclei_CP_Intensity_MedianIntensity_EdU" %in% colnames(l3)}
```

<br>

```{r rawEdUBoxplots, fig.height=5,  eval="Nuclei_CP_Intensity_MedianIntensity_EdU" %in% colnames(l3)}

p <- ggplot(l3, aes(x=factor(Barcode), y=log2(Nuclei_CP_Intensity_MedianIntensity_EdU), colour=Well))+
  geom_boxplot(outlier.size = .5)+
  labs(x="Barcode",y="Median Intensity EdU (log2)",colour="Well", title="Median EdU Intensity per Spot by Well and Barcode")+
  theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5, size=rel(1)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.8)))
p <- p + geom_hline(yintercept = log2(median(l3$Nuclei_CP_Intensity_MedianIntensity_EdU)), colour="blue", alpha=.5)
print(p)

```

```{r rawEdUBoxplotsbyWell, fig.height=5,  eval="Nuclei_CP_Intensity_MedianIntensity_EdU" %in% colnames(l3)}

p <- ggplot(l3, aes(x=factor(Well), y=log2(Nuclei_CP_Intensity_MedianIntensity_EdU), colour=factor(Barcode)))+
  geom_boxplot(outlier.size = .5)+
  labs(x="Well",y="Median Intensity EdU (log2)",colour="Barcode", title="Median EdU Intensity per Spot by Well and Barcode")+xlab("")+ylab("Median Intensity EdU (log2)")+
  theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5, size=rel(1)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.8)))
p <- p + geom_hline(yintercept = log2(median(l3$Nuclei_CP_Intensity_MedianIntensity_EdU)), colour="blue", alpha=.5)
print(p)

```


```{r rawEdUBoxplotsbyLigand, fig.height=5, eval="Nuclei_PA_Gated_EdUPositiveProportion" %in% colnames(l3)}

p <- ggplot(l3, aes(x=factor(Ligand), y=Nuclei_PA_Gated_EdUPositiveProportion, colour=factor(Barcode)))+
  geom_boxplot(outlier.size = .5)+
  coord_cartesian(y=c(0,1))+
  labs(x="Ligand",y="EdU+ Proportion", colour="Barcode", title="EdU+ Proportion per Spot by Ligand and Barcode")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust=0.5, size=rel(1)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.8)))
p <- p + geom_hline(yintercept = median(l3$Nuclei_PA_Gated_EdUPositiveProportion), colour="blue", alpha=.5)
print(p)

```



```{r, child='MEP-LINCS_Krt5Krt19Text.Rmd', eval=length(setdiff(c("Cytoplasm_CP_Intensity_MedianIntensity_KRT5","Cytoplasm_CP_Intensity_MedianIntensity_KRT19"),colnames(l3)))==0}
```

<br>



```{r rawCaspaseBoxplots, fig.height=5, eval="Cytoplasm_CP_Intensity_MedianIntensity_Caspase" %in% colnames(l3)}

p <- ggplot(l3, aes(x=factor(Barcode), y=log2(Cytoplasm_CP_Intensity_MedianIntensity_Caspase), colour=Well))+
  geom_boxplot(outlier.size = .5)+
  xlab("")+ylab("Median Intensity Caspase (log2)")+
  ggtitle("Median Intensity Caspase per Spot by Well and Barcode ")+
  theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5, size=rel(1)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.8)))
p <- p + geom_hline(yintercept = median(l3$Cytoplasm_CP_Intensity_MedianIntensity_CaspaseLog2), colour="blue", alpha=.5)
print(p)

```
<br>



```{r rawKrt5Boxplots, fig.height=5, eval="Cytoplasm_CP_Intensity_MedianIntensity_KRT5" %in% colnames(l3)}

p <- ggplot(l3, aes(x=factor(Barcode), y=log2(Cytoplasm_CP_Intensity_MedianIntensity_KRT5), colour=Well))+
  geom_boxplot(outlier.size = .5)+
  xlab("")+ylab("Median Intensity KRT5 (log2)")+
  ggtitle("Median Intensity KRT5 per Spot by Well and Barcode ")+
  theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5, size=rel(1)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.8)))
p <- p + geom_hline(yintercept = median(log2(l3$Cytoplasm_CP_Intensity_MedianIntensity_KRT5)), colour="blue", alpha=.5)
print(p)

```

```{r, child='MEP-LINCS_Krt14Krt19Text.Rmd', eval=grepl("SS6",StainingSet)}
```

<br>

```{r rawKrt14Boxplots, fig.height=5, eval="Cytoplasm_CP_Intensity_MedianIntensity_KRT14" %in% colnames(l3)}

p <- ggplot(l3, aes(x=factor(Barcode), y=log2(Cytoplasm_CP_Intensity_MedianIntensity_KRT14), colour=Well))+
  geom_boxplot(outlier.size = .5)+
  xlab("")+ylab("Median Intensity KRT14 (log2)")+
  ggtitle("Median Intensity KRT14 per Spot by Well and Barcode ")+
  theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5, size=rel(1)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.8)))
p <- p + geom_hline(yintercept = median(log2(l3$Cytoplasm_CP_Intensity_MedianIntensity_KRT14)), colour="blue", alpha=.5)
print(p)

```

<br>


```{r rawKrt19Boxplots, fig.height=5, eval="Cytoplasm_CP_Intensity_MedianIntensity_KRT19" %in% colnames(l3)}

p <- ggplot(l3, aes(x=factor(Barcode), y=log2(Cytoplasm_CP_Intensity_MedianIntensity_KRT19), colour=Well))+
  geom_boxplot(outlier.size = .5)+
  xlab("")+ylab("Median Intensity KRT19 (log2)")+
  ggtitle("Median Intensity KRT19 per Spot by Well and Barcode ")+
  theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5, size=rel(1)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.8)))
p <- p + geom_hline(yintercept = median(log2(l3$Cytoplasm_CP_Intensity_MedianIntensity_KRT19)), colour="blue", alpha=.5)
print(p)

```


<br>

```{r Filtering}
#Remove the fiducial and blank data
setkey(l3,ECMp)
l3F <- l3[!grepl("fiducial|Fiducial|blank|PBS",l3$ECMp),]

#Remove failed QA wells
l3F <- l3F[!l3F$QA_LowWellQA]
l3F <- l3F[l3F$QAScore>.6]
#Remove plate labels from FBS replicates
l3FPrep <- l3F[,grep("URL",colnames(l3F),value=TRUE, invert=TRUE), with=FALSE]
l3FPrep$Ligand <- gsub("_P[[:digit:]]","",l3FPrep$Ligand)
l3FPrep$MEP <- gsub("_P[[:digit:]]","",l3FPrep$MEP)
l3FPrep$MEP_Drug <- gsub("_P[[:digit:]]","",l3FPrep$MEP_Drug)

l4F <- preprocessLevel4(l3FPrep,seNames=c("DNA2N","SpotCellCount","EdU","MitoTracker","KRT","Fibrillarin","Caspase"))
#Add in the barcodes for each MEP_Drug
l4F <- addBarcodes(dt3 = l3FPrep, dt4 = l4F)
#Shorten Barcode to "Multiple"" if the data comes from more than one plate
l4F$Barcode[grepl(",",l4F$Barcode)] <- "Multiple"
```

```{r, child='MEP-LINCS_RUVText.Rmd', eval=unique(l3$NormMethod)=="RUV"}
```

```{r, child='MEP-LINCS_RUVLoessResidualsText.Rmd', eval=unique(l3$NormMethod)=="RUVLoessResiduals"}
```


###Spot Cell Count QA
The following figures show the raw and then normalized spot cell count values organized first by barcode and then by ligand. The barcode is replaced by the label "Multiple" when the values come from replicates that are in more than one plate.  


```{r SCC_responseFull, fig.width=8,  fig.height=10, echo=FALSE, eval="Spot_PA_SpotCellCountNorm" %in% colnames(l4F)}
dt <- l4F
dt <- l4F[,list(Barcode,Ligand,Spot_PA_SpotCellCount, Spot_PA_SpotCellCountNorm)]
dt <- dt[,Spot_PA_SpotCellCountLog2:= log2(Spot_PA_SpotCellCount)]
dt <- melt(dt,id.vars=c("Barcode","Ligand"),measure.vars=c("Spot_PA_SpotCellCount","Spot_PA_SpotCellCountNorm"), variable.name = "Processed",value.name = "SpotCellCount", variable.factor = FALSE)
dt$Processed[!grepl("Norm",dt$Processed)] <- "Raw"
dt$Processed[grepl("Norm",dt$Processed)] <- "Normalized"
dt$Processed <- factor(dt$Processed,levels = c("Raw","Normalized"))

p <- ggplot(dt, aes(x = reorder(Barcode, SpotCellCount, FUN=median), y = SpotCellCount, colour = Ligand))+geom_boxplot()+
  ggtitle(paste("MEP Normalized Spot Cell Count by Barcode"))+
  xlab("")+ylab("")+
  facet_wrap(~Processed, ncol=1)+
  guides(col = guide_legend(ncol = 2))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.5)),legend.title=element_text(size = rel(.8)))
p <- p+geom_hline(yintercept = median(dt$SpotCellCount), colour="blue", alpha=.5)
suppressWarnings(print(p))


p <- ggplot(dt, aes(x = reorder(Ligand, SpotCellCount, FUN=median), y = SpotCellCount, colour = Barcode))+geom_boxplot()+
  ggtitle(paste("MEP Normalized Spot Cell Count by Ligand"))+
  xlab("")+ylab("")+
  facet_wrap(~Processed, ncol=1)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.5)),legend.title=element_text(size = rel(.8)))
p <- p+geom_hline(yintercept = median(dt$SpotCellCount), colour="blue", alpha=.5)
suppressWarnings(print(p))
```


```{r SCC_response, fig.width=8,  fig.height=6}
setkey(l4F,"Ligand")

if (any(grepl("Spot_PA_SpotCellCountNorm",colnames(l4F)))){
  p <- ggplot(l4F, aes(x = reorder(Ligand, Spot_PA_SpotCellCountNorm, FUN=median), y = Spot_PA_SpotCellCountNorm, colour = Barcode))+geom_boxplot()+
    ggtitle(paste("\n\nMEP Normalized Spot Cell Count by Ligand"))+
    xlab("")+ylab("")+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
} else {
  p <- ggplot(l4F, aes(x = reorder(Ligand, Spot_PA_SpotCellCount, FUN=median), y = Spot_PA_SpotCellCount, colour = Barcode))+
    geom_boxplot()+
    ggtitle(paste("\n\nMEP Raw Spot Cell Count by Ligand"))+
    xlab("")+ylab("")+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
}
suppressWarnings(print(p))
```


```{r, fig.width=8,  fig.height=3.5}
dt <- l4F
if (any(grepl("Spot_PA_SpotCellCountNorm",colnames(l4F)))){
  p <- ggplot(dt, aes(x = reorder(ECMp, Spot_PA_SpotCellCountNorm, FUN=median), y = Spot_PA_SpotCellCountNorm))+geom_boxplot()+
    ggtitle(paste("\n\nMEP Normalized Spot Cell Count by ECM Protein"))+
    xlab("")+ylab("")+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
} else {
  p <- ggplot(dt, aes(x = reorder(ECMp, Spot_PA_SpotCellCount, FUN=median), y = Spot_PA_SpotCellCount))+geom_boxplot()+
    ggtitle(paste("MEP Raw Spot Cell Count by ECM Protein"))+
    xlab("")+ylab("")+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
}
suppressWarnings(print(p))

```



```{r MitoTrackerHistogram, eval=grepl("SS1",StainingSet)}

setkey(l4,"Ligand")
p <- ggplot(dt, aes(x=Cytoplasm_CP_Intensity_MedianIntensity_MitoTrackerNorm))+geom_histogram(binwidth = .01)+
  #geom_vline(xintercept = c(-2,2), colour = "blue")+
  ggtitle(paste("\n\n","MEP Normalized, MitoTracker Intensity Distribution"))+
  ylab("Count")+xlab("Normalized, MitoTracker Intensity")+
  theme(strip.text = element_text(size = 5))
suppressWarnings(print(p))
```

```{r MitoTrackerByLigand, eval=grepl("SS1",StainingSet)}

p <- ggplot(dt, aes(x = reorder(Ligand, Cytoplasm_CP_Intensity_MedianIntensity_MitoTrackerNorm, FUN=median), y = Cytoplasm_CP_Intensity_MedianIntensity_MitoTrackerNorm, colour = Barcode))+geom_boxplot()+
  ggtitle(paste("\n\nMEP Normalized, MitoTracker Intensity by Ligand"))+
  xlab("")+ylab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
suppressWarnings(print(p))

```

```{r MitoTrackerByECMp, eval=grepl("SS1",StainingSet)}

p <- ggplot(dt, aes(x = reorder(ECMp, Cytoplasm_CP_Intensity_MedianIntensity_MitoTrackerNorm, FUN=median), y = Cytoplasm_CP_Intensity_MedianIntensity_MitoTrackerNorm))+geom_boxplot()+
  ggtitle(paste("\n\nMEP Normalized, MitoTracker Intensity by ECM Protein"))+
  xlab("")+ylab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
suppressWarnings(print(p))

```


```{r CaspaseHistogram, eval=grepl("SSD",StainingSet)}

setkey(l4,"Ligand")
p <- ggplot(l4, aes(x=Cytoplasm_CP_Intensity_MedianIntensity_CaspaseNorm))+geom_histogram(binwidth = .01)+
  #geom_vline(xintercept = c(-2,2), colour = "blue")+
  ggtitle(paste("\n\n","MEP Normalized, Caspase Intensity Distribution"))+
  ylab("Count")+xlab("Raw, Caspase Intensity")+
  theme(strip.text = element_text(size = 5))
suppressWarnings(print(p))
```

```{r CaspaseByLigand, eval=grepl("SSD",StainingSet)}

p <- ggplot(dt, aes(x = reorder(Ligand, Cytoplasm_CP_Intensity_MedianIntensity_CaspaseNorm, FUN=median), y = Cytoplasm_CP_Intensity_MedianIntensity_CaspaseNorm, colour = Barcode))+geom_boxplot()+
  ggtitle(paste("MEP Normalized Caspase Intensity by Ligand"))+
  xlab("")+ylab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
suppressWarnings(print(p))

```

```{r CaspaseByECMp, eval=grepl("SSD",StainingSet)}

p <- ggplot(dt, aes(x = reorder(ECMp, Cytoplasm_CP_Intensity_MedianIntensity_CaspaseNorm, FUN=median), y = Cytoplasm_CP_Intensity_MedianIntensity_CaspaseNorm))+geom_boxplot()+
  ggtitle(paste("MEP Normalized Caspase Intensity by ECM Protein"))+
  xlab("")+ylab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
suppressWarnings(print(p))

```


```{r, child='MEP-LINCS_ProliferationText.Rmd', eval=grepl("SS2|SS4|SS6|SSA",StainingSet)}
```


```{r ProliferationHistogramFull, eval="Nuclei_PA_Gated_EdUPositiveProportionNorm" %in% colnames(l3)}

dt <- l4F

p <- ggplot(dt, aes(x=Nuclei_PA_Gated_EdUPositiveProportionNorm))+geom_histogram(binwidth = .001)+
  ggtitle(paste("\n\n","MEP Normalized, EdU+ Proportion Distribution"))+
  ylab("Count")+xlab("EdU+ Proportion (Normalized)")+
  theme(strip.text = element_text(size = 5))
suppressWarnings(print(p))
```

```{r ProliferationRawHistogram, eval="Nuclei_PA_Gated_EdUPositiveProportion" %in% colnames(l3)}

dt <- l4F

p <- ggplot(dt, aes(x=Nuclei_PA_Gated_EdUPositiveProportion))+geom_histogram(binwidth = .001)+
  ggtitle(paste("MEP EdU+ Proportion Distribution"))+
  ylab("Count")+xlab("EdU+ Proportion")+
  theme(strip.text = element_text(size = 5))
suppressWarnings(print(p))
```


```{r ProliferationByLigandFull, , eval=any(grepl("Nuclei_PA_Gated_EdUPositiveProportionNorm",colnames(dt)))}

p <- ggplot(dt, aes(x = reorder(Ligand, Nuclei_PA_Gated_EdUPositiveProportionNorm, FUN=median), y = Nuclei_PA_Gated_EdUPositiveProportionNorm, colour = Barcode))+geom_boxplot()+
  coord_cartesian(ylim = c(0,1))+
  ggtitle(paste("MEP Normalized, EdU+ Proportion by Ligand"))+
  xlab("")+ylab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
suppressWarnings(print(p))

```

```{r ProliferationByECMpFull, eval=any(grepl("Nuclei_PA_Gated_EdUPositiveProportionNorm",colnames(dt)))}

p <- ggplot(dt, aes(x = reorder(ECMp, Nuclei_PA_Gated_EdUPositiveProportionNorm, FUN=median), y = Nuclei_PA_Gated_EdUPositiveProportionNorm))+geom_boxplot()+
    coord_cartesian(ylim = c(0,1))+
  ggtitle(paste("MEP Normalized, EdU+ Proportion by ECM Protein"))+
  xlab("")+ylab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
suppressWarnings(print(p))

```



```{r, child='MEP-LINCS_LineageText.Rmd',eval=grepl("SS3|SS4",StainingSet)}
```

```{r lineageRatioHistogram, eval=any(grepl("Cytoplasm_PA_Intensity_LineageRatioNorm",colnames(dt)))}

p <- ggplot(dt, aes(x=Cytoplasm_PA_Intensity_LineageRatioNorm))+geom_histogram(binwidth = .02)+
  ggtitle(paste("MEP Normalized Lineage Ratio Distribution"))+
  ylab("Count")+xlab("Normalized Lineage Ratio")+
  theme(strip.text = element_text(size = 5))
suppressWarnings(print(p))
```

```{r LineageRatioByLigand, eval=any(grepl("Cytoplasm_PA_Intensity_LineageRatioNorm",colnames(dt)))}

p <- ggplot(dt, aes(x = reorder(Ligand, Cytoplasm_PA_Intensity_LineageRatioNorm, FUN=median), y = Cytoplasm_PA_Intensity_LineageRatioNorm, colour = Barcode))+geom_boxplot()+
  ggtitle(paste("\n\nMEP Normalized Lineage Ratio by Ligand"))+
  xlab("")+ylab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
suppressWarnings(print(p))

```

```{r lineageRatioByECMp, eval=any(grepl("Cytoplasm_PA_Intensity_LineageRatioNorm",colnames(dt)))}

p <- ggplot(dt, aes(x = reorder(ECMp, Cytoplasm_PA_Intensity_LineageRatioNorm, FUN=median), y = Cytoplasm_PA_Intensity_LineageRatioNorm))+geom_boxplot()+
  ggtitle(paste("MEP Normalized LineageRatio by ECM Protein"))+
  xlab("")+ylab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
suppressWarnings(print(p))

```


```{r KRT5ByECMp, eval=any(grepl("Cytoplasm_CP_Intensity_MedianIntensity_KRT5Norm",colnames(dt)))}

p <- ggplot(dt, aes(x = reorder(ECMp, Cytoplasm_CP_Intensity_MedianIntensity_KRT5Norm, FUN=median), y = Cytoplasm_CP_Intensity_MedianIntensity_KRT5Norm))+geom_boxplot()+
  ggtitle(paste("MEP Normalized KRT5 by ECM Protein"))+
  xlab("")+ylab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
suppressWarnings(print(p))

```

###DNA Content Analysis

All cells are stained with DAPI and classified as DNA 2n or 4n. The proportion of 2n and 4n cells at each spot is calculated and will always sum to 1. The proportions are plotted below and detailed in the datatables.


```{r DNAHistogramFull}

dt <- l3

p <- ggplot(dt, aes(x=Nuclei_PA_Cycle_DNA2NProportion))+geom_histogram(binwidth = .02, alpha=.8)+
  ggtitle(paste("Spot DNA 2n Proportion Histogam"))+
  ylab("Count")+xlab("DNA 2n Proportion")+
  theme(strip.text = element_text(size = 5))
suppressWarnings(print(p))
```

###PCA Analysis
A Principal Component Analysis can be used to evaluate the normalization process for plate and well effects.  

```{r, eval=FALSE}
###PCA Unbiased Analysis  
PCFVNames <- grep("Norm|Loess",colnames(l4F),value=TRUE)
PCFVNames <- grep("Euler|MaximumRadius|_SE|_RobustZ",PCFVNames,invert = TRUE, value = TRUE)

#Get the PCFVs
l4PCAModel <- prcomp(as.formula(paste(" ~ ", paste(PCFVNames, collapse = "+"))), data = l4F, scale= TRUE)

#Extract the first PCs of each MEP into a matrix
l4pcvDT <- data.table(l4PCAModel$x[,1:nrPCs],Barcode =l4F$Barcode, MEP = l4F$MEP)

#Convert to dataframe for use in D3heatmap
l4pcvDT <- data.frame(l4pcvDT)
rownames(l4pcvDT) <- l4pcvDT$MEP
l4pcvDT <- l4pcvDT[,!(names(l4pcvDT) %in% c("MEP", "Barcode"))]

```

```{r PCAHeatmap, fig.height=8, fig.width=7, eval=FALSE}
#Plot the heat map of the entire dataset
d3heatmap(l4pcvDT, scale="column", colors=hmcols, xaxis_font_size="6pt", yaxis_font_size="4pt", show_grid = FALSE)
```

```{r, fig.height=3, fig.width=4, eval=FALSE}

var <- l4PCAModel$sd[1:nrPCs]^2
var.percent <- var/sum(var) * 100
barplot(var.percent, xlab="PC", ylab="Percent Variance", names.arg=1:length(var.percent), las=1, ylim=c(0,max(var.percent)), col="gray", main  = paste("PCA Scree Plot"))

p <- ggplot(data.frame(l4PCAModel$x), aes(x = PC1, y = PC2, col = l4F$Barcode)) +
  geom_point(size = rel(.7), alpha = .8) +
  labs(colour = "Barcode")+
  theme(legend.text=element_text(size = 6))+
  guides(colour = guide_legend(override.aes = list(size=6)))
print(p)

p <- ggplot(data.frame(l4PCAModel$x), aes(x = PC1, y = PC3, col = l4F$Barcode)) +
  geom_point(size = rel(.7), alpha = .8) +
  labs(colour = "Barcode")+
  theme(legend.text=element_text(size = 6))+
  guides(colour = guide_legend(override.aes = list(size=6)))
print(p)

p <- ggplot(data.frame(l4PCAModel$x), aes(x = PC2, y = PC3, col = l4F$Barcode)) +
  geom_point(size = rel(.7), alpha = .8) +
  labs(colour = "Barcode")+
  theme(legend.text=element_text(size = 6))+
  guides(colour = guide_legend(override.aes = list(size=6)))
print(p)

p <- ggplot(data.frame(l4PCAModel$x), aes(x = PC1, y = PC2, col = l4F$Well)) +
  geom_point(size = rel(.7), alpha = .8) +
  labs(colour = "Well")+
  theme(legend.text=element_text(size = 6))+
  guides(colour = guide_legend(override.aes = list(size=6)))
print(p)

p <- ggplot(data.frame(l4PCAModel$x), aes(x = PC1, y = PC3, col = l4F$Well)) +
  geom_point(size = rel(.7), alpha = .8) +
  labs(colour = "Well")+
  theme(legend.text=element_text(size = 6))+
  guides(colour = guide_legend(override.aes = list(size=6)))
print(p)

p <- ggplot(data.frame(l4PCAModel$x), aes(x = PC2, y = PC3, col = l4F$Well)) +
  geom_point(size = rel(.7), alpha = .8) +
  labs(colour = "Well")+
  theme(legend.text=element_text(size = 6))+
  guides(colour = guide_legend(override.aes = list(size=6)))
print(p)

if(grepl("SS1",StainingSet)){
  p <- ggplot(data.frame(l4PCAModel$x), aes(x = PC1, y = PC2, col = l4F$Cytoplasm_CP_Intensity_IntegratedIntensity_MitoTrackerNorm)) +
    geom_point(size = rel(.7), alpha = .8) +
    labs(colour = "Total MitoTracker")+
    theme(legend.text=element_text(size = 6))+
    guides(colour = guide_legend(override.aes = list(size=6)))
  print(p)
  
  
  p <- ggplot(data.frame(l4PCAModel$x), aes(x = PC1, y = PC2, col = l4F$Cells_CP_AreaShape_PerimeterNorm)) +
    geom_point(size = rel(.7), alpha = .8) +
    labs(colour = "Perimeter ProportionStatus")+
    theme(legend.text=element_text(size = 6))+
    guides(colour = guide_legend(override.aes = list(size=6)))
  print(p)
} else if (ss=="SS2|SSA"){
  
  p <- ggplot(data.frame(l4PCAModel$x), aes(x = PC1, y = PC2, col = l4F$Nuclei_CP_Intensity_MedianIntensity_EdURUVLoess)) +
    geom_point(size = rel(.7), alpha = .8) +
    labs(colour = "Median Intensity EdU")+
    theme(legend.text=element_text(size = 6))+
    guides(colour = guide_legend(override.aes = list(size=6)))
  
} else if (ss=="SS3|SS4"){
  
  
  p <- ggplot(data.frame(l4PCAModel$x), aes(x = PC1, y = PC2, col = l4F$Cytoplasm_PA_Intensity_LineageRatioNorm)) +
    geom_point(size = rel(.7), alpha = .8) +
    labs(colour = "Lineage Ratio")+
    theme(legend.text=element_text(size = 6))+
    guides(colour = guide_legend(override.aes = list(size=6)))
  print(p)
  
  
  p <- ggplot(data.frame(l4PCAModel$x), aes(x = PC3, y = PC4, col = l4F$Cytoplasm_PA_Intensity_LineageRatioNorm)) +
    geom_point(size = rel(.7), alpha = .8) +
    labs(colour = "Lineage Ratio")+
    theme(legend.text=element_text(size = 6))+
    guides(colour = guide_legend(override.aes = list(size=6)))
  print(p)
  
  p <- ggplot(data.frame(l4PCAModel$x), aes(x = PC5, y = PC6, col = l4F$Cytoplasm_PA_Intensity_LineageRatioNorm)) +
    geom_point(size = rel(.7), alpha = .8) +
    labs(colour = "Lineage Ratio")+
    theme(legend.text=element_text(size = 6))+
    guides(colour = guide_legend(override.aes = list(size=6)))
  print(p)
  
  p <- ggplot(data.frame(l4PCAModel$x), aes(x = PC7, y = PC8, col = l4F$Cytoplasm_PA_Intensity_LineageRatioNorm)) +
    geom_point(size = rel(.7), alpha = .8) +
    labs(colour = "Lineage Ratio")+
    theme(legend.text=element_text(size = 6))+
    guides(colour = guide_legend(override.aes = list(size=6)))
  print(p)
  
}
```

```{r PCADatatable, cache=FALSE, eval=FALSE}
datatable(format(l4PCAModel$rotation[,1:nrPCs], digits=1, scientific = FALSE, nsmall=0), options = list(pageLength = 5))
#Extract the selectedPCs of each MEP into a matrix

```



```{r, eval=FALSE}


###PCA Unbiased Analysis  on raw l3 data


PCFVNames <- grep("Norm|Loess",colnames(l3F),value=TRUE)
PCFVNames <- grep("Euler|MaximumRadius|_SE|_RobustZ",PCFVNames,invert = TRUE, value = TRUE)
PCFVNames <- gsub("RUVLoess","", PCFVNames)
tmp <- grep()

#Get the PCFVs
l4PCAModel <- prcomp(as.formula(paste(" ~ ", paste(PCFVNames, collapse = "+"))), data = l4F, scale= TRUE)

#Extract the first PCs of each MEP into a matrix
l4pcvDT <- data.table(l4PCAModel$x[,1:nrPCs],Barcode =l4F$Barcode, MEP = l4F$MEP)

#Convert to dataframe for use in D3heatmap
l4pcvDT <- data.frame(l4pcvDT)
rownames(l4pcvDT) <- l4pcvDT$MEP
l4pcvDT <- l4pcvDT[,!(names(l4pcvDT) %in% c("MEP", "Barcode"))]

```

<br>

##Supplemental Material


###MEMA Layout
All MEMAs in the experiment are in separate wells and have the same design of 48 ECM proteins spotted in 35 rows and 20 columns. The proteins are randomly assigned to spots in the top 30 rows. Rows 31-35 are replicates of rows 1-5. There is a higher number of COL1 spots (shown as black triangles) throughout the array that are control spots used to evaluate spatial variations. The upper left and bottom right corners of each MEMA are image fiducials in the 488nm channel and there are four blank spots for checking orientation in all channels.  

```{r Content Layout,echo=FALSE, message=FALSE, warnings=FALSE, fig.height=6, fig.width=5, eval=exists("l3")}

#Select the A row wells and delete the blanks
setkey(l3,Well)
DT <- unique(l3[grep("A",unique(l3$Well),value=TRUE),list(ArrayRow,ArrayColumn,ECMp)])
setkey(DT,ECMp)
DT <- DT[!grepl("blank|PBS",DT$ECMp),]

p <- ggplot(data=DT[!grepl("COL1",DT$ECMp),],aes(x = ArrayColumn, y = ArrayRow, fill=ECMp))+
  geom_point(shape=21, size = 2.7)+
  guides(fill=FALSE)+
  theme(legend.text = element_text(size = rel(.5)),legend.title=element_text(size = rel(.5)),plot.title=element_text(size = rel(.8)))+
  scale_y_reverse()+
  xlab("")+ylab("")+
  ggtitle(" \n\nLINCS MEMA A Row Layout")
p <- p+geom_point(data = DT[grepl("COL1",DT$ECMp),], shape=24, fill="black")
print(p)

```

###Replicate Count
The median replicate count is `r  sprintf("%3.2f",median(table(DT$ECMp)))` and the range of the replicate counts is `r range(table(DT$ECMp))`.  

```{r Layout Replicate Count,echo=FALSE, message=FALSE, warnings=FALSE, fig.width=6.5, fig.height=8, eval=dim(DT)[1]>0}

p <- ggplot(DT, aes(x=ECMp))+
  geom_bar(width=.8)+geom_hline(yintercept = median(table(DT$ECMp)), colour="blue")+
  scale_y_log10()+
  ggtitle(" \n\nCount of Replicate ECM Proteins In Each MEMA")+
  xlab("Printed ECM Protein")+ylab("Number of spots")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)),axis.title.x = element_text(size=rel(.8)),axis.title.y = element_text(size=rel(.8)),plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.5)))

print(p)

```

###Quality Assessment
The variance of the signal in MEMA data comes from biological and technical factors. The technical factors create regions of low cell counts per spot and uneven staining across the array. The goal of the QA pipeline is to quantify the technical factors to identify wells or plates that need to be removed from downstream processing and/or be replaced by wells from a new experiment.

The hypothesis for the MEMA QA process is that the biological signal comes from individual spots while the technical variations come from regions of low signal. A bivariate loess model can be used to quantify the number of spots in low signal regions, leading to a MEMA QA score. 
<br>

####Loess Model Explanation
The loess model of a MEMA is the mean value of a weighted version of each spot's region or neighborhood. In a 700 spot array, a loess span value of 0.1 sets the size of the neighborhood to be the nearest 70 points (within approximately 5 spots in all directions). The weights are a tricubic function of the euclidean distance between the spot being modeled and the neighborhood spots. These weights vary from 1 to 0 as distances increase from the nearest to the farthest neighbor. In other words, each spot in the model takes on the mean value of its 70 nearest neighbors with the closest neighbors having the largest impact. Therefore, the loess model is dominated by the technical regional factors as opposed to individual biological responses.

A MEMA's QA score is derived from the loess model of the control-well-normalized values by calculating the proportion of spots in low signal regions(LSR). A threshold for classifying spots as LSR is based on the median of each plate's control well. To have higher scores reflect increasing quality, the MEMA QA score is defined as the proportion of non-LSR spots to total spots. This value will be 1 for MEMAs with no low signal regions and approach 0 as the number of LSR spots increases.

In the plots below, the LSR spots are those to the left of the blue vertical line at the threshold value of `r wellQAThresh` in the histogram. 


```{r Loess_Model_explanation , echo=FALSE, fig.width=2.5,fig.height=4}
setkey(l3,Barcode,Well)
wellExample <- unique(l3$Well)[1]
DT <-l3[.(barcodes[1],wellExample)]
#Remove the fiducial entries
setkey(DT,ECMp)
DT <- DT[!"fiducial"]
DT <- DT[!"blank"]

p <- ggplot(DT, aes(x=ArrayColumn, y=ArrayRow,colour=Spot_PA_SpotCellCount))+
  geom_point(size=1.8)+
  scale_y_reverse()+   scale_x_continuous(breaks= c(min(DT$ArrayColumn),round(mean(c(min(DT$ArrayColumn),max(DT$ArrayColumn)))),max(DT$ArrayColumn)))+
  scale_colour_gradient(low = "white", high = "red")+
  guides(colour = guide_legend("Spot Cell\nCount", keywidth = .5, keyheight = .5))+
  ggtitle(paste("\n\n","Spot Cell Count for",unique(DT$CellLine), "cells \nin plate",unique(DT$Barcode)))+
  xlab("")+ylab("")+theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),plot.title = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
suppressWarnings(print(p))

p <- ggplot(DT, aes(x=ArrayColumn, y=ArrayRow,colour=Spot_PA_LoessSCC))+
  geom_point(size=1.8)+
  scale_y_reverse()+   scale_x_continuous(breaks= c(min(DT$ArrayColumn),round(mean(c(min(DT$ArrayColumn),max(DT$ArrayColumn)))),max(DT$ArrayColumn)))+
  scale_colour_gradient(low = "white", high = "red")+
  guides(colour = guide_legend("Normalized \nSpot Cell \nCount", keywidth = .5, keyheight = .5))+
  ggtitle(paste("\n\n","Loess Model of Spot Cell Count \nfor",unique(DT$CellLine), "cells in plate",unique(DT$Barcode)))+
  xlab("")+ylab("")+theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),plot.title = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
suppressWarnings(print(p))

DT <- DT[,QAScore := calcQAScore(.SD,threshold=.6,value="Spot_PA_LoessSCC"),by="Well"]
wellScores <- unique(DT[,list(Well,QAScore=sprintf("%.2f",QAScore))])

p <- ggplot(DT, aes(x=Spot_PA_LoessSCC))+
  geom_histogram(binwidth=.02)+
  geom_vline(xintercept=.6, colour="blue")+
  geom_text(data=wellScores, aes(label=paste0("QA\n",QAScore)), x = .9, y = 30, size = rel(5), colour="red")+
  ggtitle(paste("\n\n","Loess Model of Spot Cell Count \nfor",unique(DT$CellLine), "cells in plate",unique(DT$Barcode)))+xlab("Spot Cell Count")+
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
suppressWarnings(print(p))


```


###Stain Pseudo images
The pseudo images of each well's raw signals are shown in the plots below. Wells that could not be sucessfully imaged due to focus issues are missing from the pseudoimages.

```{r Pseudoimages_all_stains, echo=FALSE, fig.width=3.7,fig.height=4, eval=TRUE}

for (barcode in barcodes){
  DT <-l3[l3$Barcode==barcode,]
  #Remove the fiducial entries
  setkey(DT,ECMp)
  DT <- DT[!"fiducial"]
  DT <- DT[!"blank"]
  DT$Well_Ligand <- sub("_","\n",DT$Well_Ligand)
  
  
  p <- create8WellPseudoImage(DT, pr = "Nuclei_CP_Intensity_MedianIntensity_Dapi", prDisplay = "Median DAPI")
  suppressWarnings(print(p))
  
  
  if (any(grepl("Actin",colnames(DT))))  {
    p <- create8WellPseudoImage(DT, pr = "Cytoplasm_CP_Intensity_MedianIntensity_Actin", prDisplay = unique(DT$Endpoint488))
    suppressWarnings(print(p))
  }
  if (any(grepl("CellMask",colnames(DT))))  {
    p <- create8WellPseudoImage(DT, pr = "Cytoplasm_CP_Intensity_MedianIntensity_CellMask",prDisplay = unique(DT$Endpoint555))
    suppressWarnings(print(p))
  }
  if (any(grepl("MitoTracker",colnames(DT))))  {
    p <- create8WellPseudoImage(DT, pr = "Cytoplasm_CP_Intensity_MedianIntensity_MitoTracker",prDisplay = unique(DT$Endpoint647))
    suppressWarnings(print(p))
  }
  
  if (any(grepl("Fibrillarin",colnames(DT))))  {
    p <- create8WellPseudoImage(DT, pr = "Nuclei_CP_Intensity_MedianIntensity_Fibrillarin",prDisplay = unique(DT$Endpoint555))
    suppressWarnings(print(p))
  }
  if (any(grepl("EdU",colnames(DT))))  {    
    p <- create8WellPseudoImage(DT, pr = "Nuclei_CP_Intensity_MedianIntensity_EdU",prDisplay = unique(DT$Endpoint647))
    suppressWarnings(print(p))
  } 
  
  if (any(grepl("KRT5",colnames(DT))))  {    
    p <- create8WellPseudoImage(DT, pr = "Cytoplasm_CP_Intensity_MedianIntensity_KRT5", prDisplay = unique(DT$Endpoint488))
    suppressWarnings(print(p))
  }
  if (any(grepl("KRT19",colnames(DT))))  {
    p <- create8WellPseudoImage(DT, pr = "Cytoplasm_CP_Intensity_MedianIntensity_KRT19",prDisplay = unique(DT$Endpoint555))
    suppressWarnings(print(p))
  }
  
  if (any(grepl("KRT14",colnames(DT))))  {
    
    p <- create8WellPseudoImage(DT, pr = "Cytoplasm_CP_Intensity_MedianIntensity_KRT14", prDisplay = unique(DT$Endpoint488))
    suppressWarnings(print(p))
  }
  
  if (any(grepl("Caspase",colnames(DT))))  {    
    p <- create8WellPseudoImage(DT, pr = "Nuclei_CP_Intensity_IntegratedIntensity_Caspase", prDisplay = unique(DT$Endpoint488))
    suppressWarnings(print(p))
  }
  
  if (any(grepl("Nestin",colnames(DT))))  {    
    p <- create8WellPseudoImage(DT, pr = "Nuclei_CP_Intensity_MedianIntensity_Nestin", prDisplay = unique(DT$Endpoint488))
    suppressWarnings(print(p))
  }
  
}
```


```{r, eval=length(setdiff(c("Nuclei_CP_Intensity_IntegratedIntensity_Fibrillarin","Nuclei_CP_Intensity_MedianIntensity_EdU"),colnames(l3)))==0}
p <- ggplot(l3, aes(x=factor(Barcode), y=Nuclei_CP_Intensity_IntegratedIntensity_Fibrillarin, colour=Well))+
  geom_boxplot(outlier.size = .5)+
  coord_cartesian(ylim = c(0,20))+
  guides(colour=FALSE)+
  xlab("")+ylab("Total Intensity Fibrillarin")+
  ggtitle("Total Fibrillarin Intensity per Spot by Well and Barcode ")+
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(.8)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
print(p)    

p <- ggplot(l3, aes(x=factor(Barcode), y=Nuclei_CP_Intensity_MedianIntensity_EdU, colour=Well))+
  geom_boxplot(outlier.size = .5)+
  #coord_cartesian(ylim = c(0,.04))+
  guides(colour=FALSE)+
  xlab("")+ylab("Median Intensity EdU")+
  ggtitle("Median EdU Intensity per Spot by Well and Barcode ")+
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(.8)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
print(p)


```


```{r, eval="Nuclei_CP_Intensity_MedianIntensity_EdU" %in% colnames(l3)}


p <- ggplot(l3, aes(x=factor(Barcode), y=Nuclei_CP_Intensity_MedianIntensity_EdU, colour=Well))+
  geom_boxplot(outlier.size = .5)+
  guides(colour=FALSE)+
  xlab("")+ylab("Median Intensity EdU")+
  ggtitle("Median EdU Intensity per Spot by Well and Barcode ")+
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(.8)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
print(p)


```

```{r, , fig.height=10,  eval="Nuclei_PA_Gated_EdUPositiveProportion" %in% colnames(l3)}

  p <- ggplot(l3, aes(y=Nuclei_PA_Gated_EdUPositiveProportion, x=Well, colour=Well))+
    geom_boxplot()+
    xlab("Well")+
    coord_cartesian( ylim=c(0,1))+
    ggtitle("EdU+ Proportion by Well")+
    facet_wrap(~Barcode)+
    theme(strip.text = element_text(size = rel(.5)), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
  print(p)

```
