
```{r "setup", include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE, cache.lazy = FALSE)

library(rmarkdown)
library(tidyverse)
library(MEMA)
library(RColorBrewer)
library(knitr)

#Setup colors for Barcode and text in all heatmaps
selDark2 <- colorRampPalette(brewer.pal(8,"Dark2"))
plateCol = selDark2(8)
hmcols<-colorRampPalette(c("blue","white","red"))(16)
sampleProportion <- .5

#Command line arguments
path <- "/eppec/storage/groups/heiserlab/lincs"
studyName <- "hcc1143_low_serum"
```

```{r readLevel1Data}
#barcodes <-getBarcodes(studyName, synId = "syn10846457")
#identify barcodes for getting imageID metadata
barcodes <- c("LI8V01171", "LI8V01172", "LI8V01173a", "LI8V01174","LI8V01175", "LI8V01176", "LI8V01177", "LI8V01178", "LI8V01179")

l1 <- lapply(barcodes, function(barcode, path){
  sd <- read_csv(paste0(path,"/",barcode,"/Analysis/",barcode,"_level1.csv"))
}, path=path) %>%
  bind_rows()

#redefine barcodes based on the values in the data
barcodes <- unique(l1$Barcode)
set.seed(42)
l1S <- l1 %>%
  sample_frac(sampleProportion) %>%
  mutate(Well_Ligand )

#DEBUG DNAThresh used to create blue line at gate
# if("Nuclei_PA_Cycle_State" %in% colnames(l1S)){
#   l1S <- l1S %>%
#     group_by(Barcode) %>%
#     mutate()
#   l1S <- l1S[,TotalDNANormed := Nuclei_CP_Intensity_IntegratedIntensity_Dapi/median(Nuclei_CP_Intensity_IntegratedIntensity_Dapi[Nuclei_PA_Cycle_State==1], na.rm=TRUE), by="Barcode"]
#   l1S <- l1S[,DNAThresh := min(TotalDNANormed[Nuclei_PA_Cycle_State==2]), by="Barcode"]
# }
#Determine number of rows in the plate for figures
nrPlateRow <- length(unique(str_extract(l1S$Well,"[[:alpha:]]")))
CellLine <- unique(l1S$CellLine)

foo <- l1 %>%
  select(starts_with("Endpoint",ignore.case = FALSE)) %>%
  distinct %>%
  unlist
```



#MEP-LINCS `r studyName` Cell-Level QA Report
####date: `r Sys.Date()`

##Summary
This report covers the cell-level data for the `r studyName` dataset. There are additional reports for the spot and MEP QA and the analysis of this staining set. 

##Introduction
The LINCS `r studyName` experiment was performed with `r CellLine` cells grown in `r length(barcodes)` 8-well plates. The  channels are `r  combine_words(gsub("Endpoint","",grep("Endpoint",colnames(l1S),value=TRUE)))`. The endpoints are `r combine_words(l1 %>%
  select(starts_with("Endpoint",ignore.case = FALSE)) %>%
  distinct %>%
  unlist)`. Color images of the cells at each spot were gathered on a Nikon automated microscope.

The histograms and scatterplots that display cell level data are from a randomly sampled subset of the full dataset.

<br>

###DAPI QA
Histograms of the total DAPI signal are typicaly bimodal with peaks at the 2N and 4N DNA positions. If the DAPI signal passed the QA test, a blue vertical line shows the gate value for classifying 2N and 4N populations.

For DAPI signals that have been gated, we want to normalize the DNA signal so that the G0/G1 peak has a value of 1. This should yield G2 peak at 2. To achieve this, the total DAPI intensity is divided by the median total DAPI intensity for all cells that have a DNA Cycle State value of 1 (to the left of the blue line).

The plots below are of a randomly sampled subset of the cell-level data.

<br>

```{r rawDAPIPlots, fig.width=3.7,fig.height=1.5*nrPlateRow}

for(barcode in barcodes){
  dtb <- l1S %>%
    filter(Barcode == barcode) 
  lower_dapi_int <- quantile(dtb$Nuclei_CP_Intensity_IntegratedIntensity_Dapi, .005)
  upper_dapi_int <- quantile(dtb$Nuclei_CP_Intensity_IntegratedIntensity_Dapi, .98)
  
  dtb <- dtb %>%
    filter(Nuclei_CP_Intensity_IntegratedIntensity_Dapi >lower_dapi_int,
           Nuclei_CP_Intensity_IntegratedIntensity_Dapi< upper_dapi_int)
  
  p <- ggplot(dtb, aes(x=Nuclei_CP_Intensity_IntegratedIntensity_Dapi))+
    geom_histogram(bins = 300)+
    #coord_cartesian(x=c(100000,16460754))+
    facet_wrap(~Well_Ligand, nrow=2) +
    ylab("Count")+xlab("Total Intensity Dapi")+ggtitle(paste("Total Dapi Intensity:",barcode))+
    theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5, size=rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.8)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)),strip.text.x = element_text(size = rel(.4)))
  suppressWarnings(print(p))
}

```

```{r, child='MEP-LINCS_EdUText.Rmd', eval="Nuclei_CP_Intensity_MedianIntensity_EdU" %in% colnames(l1S)}
```

<br>

```{r rawEdUHistograms, fig.width=3.7,fig.height=4, eval="Nuclei_CP_Intensity_MedianIntensity_EdU" %in% colnames(l1S)}

rf <- colorRampPalette(rev(brewer.pal(11,'Spectral')))
r <- rf(32)

for(barcode in barcodes){
  dtb <- l1S %>%
    filter(Barcode == barcode) 
  lower_dapi_int <- quantile(dtb$Nuclei_CP_Intensity_IntegratedIntensity_Dapi, .005)
  upper_dapi_int <- quantile(dtb$Nuclei_CP_Intensity_IntegratedIntensity_Dapi, .98)
  
  dtb <- dtb %>%
    filter(Nuclei_CP_Intensity_IntegratedIntensity_Dapi >lower_dapi_int,
           Nuclei_CP_Intensity_IntegratedIntensity_Dapi< upper_dapi_int)
  
  p <- ggplot(dtb, aes(x=Nuclei_CP_Intensity_IntegratedIntensity_Dapi, y = log2(Nuclei_GT_EdU_MeanIntensity))) +
    geom_hex(size = .8, bins=100) +
    scale_fill_gradientn(colours=r, trans="log") +
    #coord_cartesian(x=c(5,7))+
    facet_wrap(~Well_Ligand, nrow=2) +
    ylab("Mean intensity EdU")+xlab("Total Intensity Dapi")+ggtitle(paste("Total Dapi Intensity vs Mean EdU:",barcode))+
    theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5, size=rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.8)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)),strip.text.x = element_text(size = rel(.4)))
  print(p)
}

for(barcode in barcodes){
  dtb <- l1S %>%
    filter(Barcode == barcode) 
  lower_dapi_int <- quantile(dtb$Nuclei_CP_Intensity_IntegratedIntensity_Dapi, .005)
  upper_dapi_int <- quantile(dtb$Nuclei_CP_Intensity_IntegratedIntensity_Dapi, .98)
  
  dtb <- dtb %>%
    filter(Nuclei_CP_Intensity_IntegratedIntensity_Dapi >lower_dapi_int,
           Nuclei_CP_Intensity_IntegratedIntensity_Dapi< upper_dapi_int)
  
  p <- ggplot(dtb, aes(x=Nuclei_CP_Intensity_IntegratedIntensity_Dapi, y = log2(Nuclei_GT_EdU_MeanIntensity), colour = Nuclei_PA_Gated_EdUPositive)) +
    geom_point(size = .05, alpha=.05) +
    #coord_cartesian(x=c(5,7))+
    guides(colour=FALSE)+
    facet_wrap(~Well_Ligand, nrow=2) +
    ylab("Mean intensity EdU")+xlab("Total Intensity Dapi")+ggtitle(paste("Total Dapi Intensity vs Mean EdU:",barcode))+
    theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5, size=rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.8)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)),strip.text.x = element_text(size = rel(.4)))
  print(p)
}

for(barcode in barcodes){
  dtb <- l1S %>%
    filter(Barcode == barcode) 
  lower_dapi_int <- quantile(dtb$Nuclei_CP_Intensity_IntegratedIntensity_Dapi, .005)
  upper_dapi_int <- quantile(dtb$Nuclei_CP_Intensity_IntegratedIntensity_Dapi, .98)
  
  dtb <- dtb %>%
    filter(Nuclei_CP_Intensity_IntegratedIntensity_Dapi >lower_dapi_int,
           Nuclei_CP_Intensity_IntegratedIntensity_Dapi< upper_dapi_int)
  
  p <- ggplot(dtb, aes(x=Nuclei_CP_Intensity_IntegratedIntensity_Dapi, y = log2(Nuclei_GT_EdU_MeanIntensity), colour = Nuclei_PA_Gated_EdUPositive)) +
    geom_point(size = .2, alpha=.2) +
    #coord_cartesian(x=c(5,7))+
    facet_wrap(~Well_Ligand, nrow=2) +
    ylab("Mean intensity EdU")+xlab("Total Intensity Dapi")+ggtitle(paste("Total Dapi Intensity vs Mean EdU:",barcode))+
    guides(colour=FALSE)+
    theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5, size=rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.8)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)),strip.text.x = element_text(size = rel(.4)))
  print(p)
}

```

<br>

```{r rawKRT14Histograms, fig.width=3.7,fig.height=4, eval="Cytoplasm_CP_Intensity_MedianIntensity_KRT14" %in% colnames(l1S)}
for(barcode in barcodes){
  dtb <- l1S %>%
    filter(Barcode == barcode) 
  lower_thresh <- quantile(dtb$Cytoplasm_CP_Intensity_MedianIntensity_KRT14, .005, na.rm = TRUE)
  upper_thresh <- quantile(dtb$Cytoplasm_CP_Intensity_MedianIntensity_KRT14, .98, na.rm = TRUE)
  
  dtb <- dtb %>%
    filter(Cytoplasm_CP_Intensity_MedianIntensity_KRT14 >lower_thresh,
           Cytoplasm_CP_Intensity_MedianIntensity_KRT14< upper_thresh)
  
  p <- ggplot(dtb, aes(x=Cytoplasm_CP_Intensity_MedianIntensity_KRT14))+
    geom_histogram(bins = 300)+
    #coord_cartesian(x=c(100000,16460754))+
    facet_wrap(~Well_Ligand, nrow=2) +
    ylab("Count")+xlab("Total Intensity KRT14")+ggtitle(paste("Total KRT14 Intensity:",barcode))+
    theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5, size=rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.8)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)),strip.text.x = element_text(size = rel(.4)))
  suppressWarnings(print(p))
}

```


<br>

```{r rawVimentinHistograms, fig.width=3.7,fig.height=4, eval="Cytoplasm_CP_Intensity_MedianIntensity_VIM" %in% colnames(l1S)}

for(barcode in barcodes){
  dtb <- l1S %>%
    filter(Barcode == barcode) 
  lower_thresh <- quantile(dtb$Cytoplasm_CP_Intensity_MedianIntensity_VIM, .005, na.rm = TRUE)
  upper_thresh <- quantile(dtb$Cytoplasm_CP_Intensity_MedianIntensity_VIM, .98, na.rm = TRUE)
  
  dtb <- dtb %>%
    filter(Cytoplasm_CP_Intensity_MedianIntensity_VIM >lower_thresh,
           Cytoplasm_CP_Intensity_MedianIntensity_VIM< upper_thresh)
  
  p <- ggplot(dtb, aes(x=Cytoplasm_CP_Intensity_MedianIntensity_VIM))+
    geom_histogram(bins = 300)+
    #coord_cartesian(x=c(100000,16460754))+
    facet_wrap(~Well_Ligand, nrow=2) +
    ylab("Count")+xlab("Total Intensity VIM")+ggtitle(paste("Total VIM Intensity:",barcode))+
    theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5, size=rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.8)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)),strip.text.x = element_text(size = rel(.4)))
  suppressWarnings(print(p))
}

```


#Supplemental Material



```{r KRT5EdU_plots,echo=FALSE, fig.width=8, fig.height=5, eval=length(setdiff(c("Cytoplasm_CP_Intensity_MedianIntensity_KRT14","Nuclei_CP_Intensity_MedianIntensity_EdU","Nuclei_PA_Gated_EdUPositive"),colnames(l1S)))==0}
binwidth=.04
for (barcode in barcodes){
  dtb <- l1S %>%
    filter(Barcode == barcode) 
  
  if(!anyNA(dtb$Nuclei_CP_Intensity_MedianIntensity_EdU)){
    p <- ggplot(dtb, aes(x=log2(Cytoplasm_CP_Intensity_MedianIntensity_KRT14), y = log2(Nuclei_CP_Intensity_MedianIntensity_EdU), colour = Nuclei_PA_Gated_EdUPositive))+geom_point(size = rel(.2), alpha = .4)+
      geom_rug(col=rgb(.5,0,0,alpha=.01))+
      scale_x_continuous(name="Median Intensity KRT14 \n(log2)",breaks=0:6)+
      facet_wrap(~Well_Ligand, nrow=2)+ggtitle(paste("EdU vs. KRT5,",barcode))+ylab("EdU\n(log2)")+
      guides(colour=FALSE)+
      theme(strip.text = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
    suppressWarnings(print(p))
  }
}

```


```{r VIMEdU_plots,echo=FALSE, fig.width=8, fig.height=5, eval=length(setdiff(c("Cytoplasm_CP_Intensity_MedianIntensity_VIM","Nuclei_CP_Intensity_MedianIntensity_EdU","Nuclei_PA_Gated_EdUPositive"),colnames(l1S)))==0}
binwidth=.04
for (barcode in barcodes){
  dtb <- l1S %>%
    filter(Barcode == barcode) 
  
  if(!anyNA(dtb$Nuclei_CP_Intensity_MedianIntensity_EdU)){
    p <- ggplot(dtb, aes(x=log2(Cytoplasm_CP_Intensity_MedianIntensity_VIM), y = log2(Nuclei_CP_Intensity_MedianIntensity_EdU), colour = Nuclei_PA_Gated_EdUPositive))+geom_point(size = rel(.2), alpha = .4)+
      geom_rug(col=rgb(.5,0,0,alpha=.01))+
      scale_x_continuous(name="Median Intensity VIM \n(log2)",breaks=0:6)+
      facet_wrap(~Well_Ligand, nrow=2)+ggtitle(paste("EdU vs. KRT5,",barcode))+ylab("EdU\n(log2)")+
      guides(colour=FALSE)+
      theme(strip.text = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
    suppressWarnings(print(p))
  }
}

```

