---
output: html_document
---

---
title: "MEP-LINCS_PC3_CV_Analysis"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, 
                      echo=FALSE, warning=FALSE, message=FALSE)
```


```{r }
#Author: Mark Dane, copyright 2015

source("../MEPLINCSFunctions.R")

#Set the cell line to be analyzed (PC3|MCF7|YAPC)
cellLine <- "PC3"

permNumber <- 50 #Used to create null hypotheses distributions

```




```{r setup}
library("ggplot2")
library("data.table")
library("MEMA")
library("grid")
library(knitr)
library("gplots")
library("RColorBrewer")
library("parallel")
library("limma")


#Setup colors for Barcode and text in all heatmaps
selDark2 <- colorRampPalette(brewer.pal(8,"Dark2"))
plateCol = selDark2(8)
ligandTypeCols = selDark2(28)
selReds <- colorRampPalette(brewer.pal(9,"Reds"))
redsPal <- selReds(10)

integrateSSCommon <- function(SSs, level, cellLine = "PC3"){
  #browser()
  lList <- lapply(SSs, function(ss){
    l <- fread(paste0("./",ss,"/AnnotatedData/",cellLine,"_",ss,"_Level",level,".txt"), showProgress = FALSE)
    setkey(l,"ECMp")
    l <- l[!"fiducial"]
    l <- l[!"blank"]
    setkey(l, "MEP")
    l$SS <- ss
    return(l)
  })
  
  lSS1 <- lList[[1]]
  lSS2 <- lList[[2]]
  lSS3 <- lList[[3]]
  
  commonNames <- intersect(intersect(colnames(lSS1),colnames(lSS2)),colnames(lSS3))
  
  setkey(lSS1,"LigandAnnotID","ECMpAnnotID")
  setkey(lSS2,"LigandAnnotID","ECMpAnnotID")
  setkey(lSS3,"LigandAnnotID","ECMpAnnotID")
  
  
  
  #Bind the data using the common MEPs
  DT <- rbind(lSS1[,commonNames, with = FALSE],
              lSS2[,commonNames, with = FALSE],
              lSS3[,commonNames, with = FALSE])
}


ranECMp <- function(x){
  #browser()
  ranX <- x[sample(1:length(x), replace = FALSE)]
  return(ranX)
}
#Sourced from http://stats.stackexchange.com/questions/78849/measure-for-separability/78855#78855

separability.measures <- function ( Vector.1 , Vector.2 ) {
# convert vectors to matrices in case they are not
  Matrix.1 <- as.matrix (Vector.1)
  Matrix.2 <- as.matrix (Vector.2)
# define means
mean.Matrix.1 <- mean ( Matrix.1 )
mean.Matrix.2 <- mean ( Matrix.2 )
# define difference of means
mean.difference <- mean.Matrix.1 - mean.Matrix.2
# define covariances for supplied matrices
cv.Matrix.1 <- cov ( Matrix.1 )
cv.Matrix.2 <- cov ( Matrix.2 )
# define the halfsum of cv's as "p"
p <- ( cv.Matrix.1 + cv.Matrix.2 ) / 2
# --%<------------------------------------------------------------------------
# calculate the Bhattacharryya index
bh.distance <- 0.125 *t ( mean.difference ) * p^ ( -1 ) * mean.difference +
0.5 * log (det ( p ) / sqrt (det ( cv.Matrix.1 ) * det ( cv.Matrix.2 )
)
)
# --%<------------------------------------------------------------------------
# calculate Jeffries-Matusita
# following formula is bound between 0 and 2.0
jm.distance <- 2 * ( 1 - exp ( -bh.distance ) )
# also found in the bibliography:
# jm.distance <- 1000 * sqrt (   2 * ( 1 - exp ( -bh.distance ) )   )
# the latter formula is bound between 0 and 1414.0
# --%<------------------------------------------------------------------------
# calculate the divergence
# trace (is the sum of the diagonal elements) of a square matrix
trace.of.matrix <- function ( SquareMatrix ) {
sum ( diag ( SquareMatrix ) ) }
# term 1
divergence.term.1 <- 1/2 * trace.of.matrix (( cv.Matrix.1 - cv.Matrix.2 ) * 
( cv.Matrix.2^ (-1) - cv.Matrix.1^ (-1) )
)
# term 2
divergence.term.2 <- 1/2 * trace.of.matrix (( cv.Matrix.1^ (-1) + cv.Matrix.2^ (-1) ) *
( mean.Matrix.1 - mean.Matrix.2 ) *
t ( mean.Matrix.1 - mean.Matrix.2 )
)
# divergence
divergence <- divergence.term.1 + divergence.term.2
# --%<------------------------------------------------------------------------
# and the transformed divergence
transformed.divergence  <- 2 * ( 1 - exp ( - ( divergence / 8 ) ) )
indices <- data.frame(
jm=jm.distance,bh=bh.distance,div=divergence,tdiv=transformed.divergence)
return(indices)
}

# Calculate Z prime factor
calculateZPrime<-function(pos,neg){
  #Given a vector of negative control values and positive control values, calculate
  #the Z Prime Factor value
  1-3*(sd(pos, na.rm = TRUE)+sd(neg, na.rm = TRUE))/abs(mean(pos, na.rm = TRUE)-mean(neg, na.rm = TRUE))
}

```


```{r, fig.height=16, eval=TRUE}

#Start with the full combined dataset with raw, cell level data
l1 <- integrateSSCommon(SSs = c("SS1","SS2","SS3"), level = 1, cellLine = "PC3")

DT1 <- copy(l1)

# selectedMEPs <- unique(DT1$LigandAnnotID)[sample(1:length(unique(DT1$MEP)), 200, replace = FALSE)]
# DT1 <- DT1[DT1$MEP %in% selectedMEPs]

DT1 <- DT1[DT1$Well =="A02"]

# #Keep only the FBS wells
# setkey(DT1,"Ligand")
# DT1 <- DT1["FBS"]

#Calculate the coefficients of variation for selected parameters
numericParameters <- colnames(DT1)[unlist(lapply(DT1,class)) %in% c("numeric","integer")]

#Remove these parameters
CVParameters <- grep("Center|Norm|_PA_|_SE|Euler|Orientation|Solidity",grep("Nuclei",numericParameters, value = TRUE),value = TRUE, invert = TRUE)

#Keep these metadata parameters
metadataNames <- grep("Barcode|^Well$|^Spot$|MEP|LigandAnnotID",colnames(DT1), value = TRUE)
DT1 <- cbind(DT1[,metadataNames, with = FALSE],DT1[,CVParameters, with = FALSE])

l3<-DT1[,lapply(.SD,numericMedian),keyby="Barcode,Well,Spot,MEP,LigandAnnotID", .SDcols = CVParameters]

CVDT <- l3[,lapply(.SD, CV), by="Barcode,Well,MEP,LigandAnnotID", .SDcols = CVParameters]
CVDT$Order <- "Actual"

#Create a null hypothesis dataset by randomizing the values within each well
#Create many random datasets to get a true null hypothesis
ranCVDTList <- mclapply(1:permNumber, function(x, dt){
  set.seed(x)
  dtRandom <- dt[,lapply(.SD, randomizePositions), by="Barcode,LigandAnnotID,Well", .SDcols = CVParameters]
  dtRandom$MEP <- dt$MEP
  dtRandom$Spot <- dt$Spot
  
  ranCVDT <- dtRandom[,lapply(.SD, CV), by="Barcode,Well,MEP,LigandAnnotID", .SDcols = CVParameters]
  ranCVDT$Order <- "Random"
  return(ranCVDT)
}, dt = l3, mc.cores = 4)

ranCVDT <- rbindlist(ranCVDTList)
CVDT <- rbind(CVDT,ranCVDT)

featureVector <- CVParameters

```

##Dataset
This data includes a curated set of the image analysis features that are common across the three PC3 staining sets. Feature classes are DAPI intensities and nuclear morphologies.

###CV Exploration
The A03 wells in each plate have replicate MEMAs containing 46 identically arranged printed ECM proteins paired with FBS in solution. Coefficients of Variation (CVs) are calculated on the feature vectors of each set of ~15 replicates in each well and on `r permNumber` permutations that randomize the assignments of ECM proteins to the same well data to create null hypotheses datasets. 

Density plots of the CVs show the experimental data in black and the null hypotheses distributions in blue. Features with actual CV distributions that are to the left of the null hypothesis show evidence of good signals. Features with CV distribitions that overlap the null hypothesis are poor quality signals. 

An hypothesis for a difference in signal quality is that the variance in the identification of the nuclear boundary is of the same magnitude as the response induced by the ECM proteins. This variance is ~squared in signals that include area (integrated intensity, area,…) and is reduced in ‘shape’ signals (eccentricity, extent,…) that include ratios.

```{r, eval=TRUE, fig.height=100}

for(feature in featureVector){
  titleName <- gsub("NA","", paste(strsplit2(feature, split = "_")[4], strsplit2(feature, split = "_")[5]))
  p <-  ggplot(CVDT[CVDT$Order == "Actual"], aes_string(x = feature))+
    geom_density(fill = "black",binwidth = .02, alpha = .5)+
    geom_density(data = CVDT[CVDT$Order == "Random"], aes_string(x = feature), fill = "blue", binwidth = .02, alpha = .5)+
    xlim(0,1)+
    facet_wrap(~MEP, ncol=8, scales="free_y")+
    ggtitle(titleName)+
    theme(strip.text = element_text(size = rel(.3)))
  print(p)
}

```

##Plate Heat Maps
The following are plate heat maps of selected features. The higher quality signals such as eccentricity show less spatial variation.


```{r, fig.height=20,fig.width=12, eval=FALSE}
curatedFeatureVector <- grep("Eccentricity|Area^|MedianIntensity|IntegratedIntensity|SpotCellCount",featureVector, value=TRUE)
for(feature in curatedFeatureVector){
  titleName <- gsub("NA","", paste(strsplit2(feature, split = "_")[4], strsplit2(feature, split = "_")[5]))
  
  p <-  ggplot(DT, aes_string(x = "ArrayColumn", y = "ArrayRow", color = feature))+
    geom_point(size =3)+
    xlim(1,20)+ylim(35,1)+
    scale_color_gradient(low="white",high="red")+
    facet_wrap(~Barcode, ncol=4)+
    ggtitle(paste(titleName, "in Control Wells"))
  print(p)
}

```

##Plate Level Boxplots
The following boxplots show the curated feature vectors stratified by plate. Each boxplot summarizes every spot in a plate. Observations include high variance in the area and spot cell count signals and lower variance in the eccentricity and median intensity signal. There are also obvious staining set effects and high variance plate effects in the DAPI signals of plates 411,413, 416, 426 and 427.


```{r, eval=FALSE}

for(feature in curatedFeatureVector){
  titleName <- gsub("NA","", paste(strsplit2(feature, split = "_")[4], strsplit2(feature, split = "_")[5]))
  
  p <-  ggplot(DT, aes_string(x="Barcode", y = feature, colour = "SS"))+
    geom_boxplot()+
    ggtitle(titleName)+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)),axis.title.x = element_text(size=rel(.8)),axis.title.y = element_text(size=rel(.8)),plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.5)))
  print(p)
}

```


##Plate Level Boxplots by ECM Protein
The following boxplots show the individual ECM protein responses of selected feature vectors stratified by plate.

```{r, fig.height=20, eval=FALSE}
p <-  ggplot(DT, aes(x=reorder(ECMpAnnotID, Nuclei_CP_AreaShape_Eccentricity_RZSNorm, FUN=median), y = Nuclei_CP_AreaShape_Eccentricity_RZSNorm, colour = SS))+
  geom_boxplot()+
  facet_wrap(~Barcode, ncol=4)+
  xlab("")+ylab("Eccentricity")+
  ggtitle("Eccentricity In Control Wells")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)),axis.title.x = element_text(size=rel(.8)),axis.title.y = element_text(size=rel(.8)),plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.5)))
print(p)
```



```{r, fig.height=20,fig.width=12, eval=FALSE}

p <-  ggplot(DT, aes(x=reorder(ECMpAnnotID,Nuclei_CP_AreaShape_Area_RZSNorm,FUN = median), y = Nuclei_CP_AreaShape_Area_RZSNorm, colour = SS))+
  geom_boxplot()+
  #xlim(1,20)+ylim(35,1)+
  #scale_color_gradient(low="white",high="red")+
  facet_wrap(~Barcode, ncol=4)+
  ggtitle("Area In Control Wells")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)),axis.title.x = element_text(size=rel(.8)),axis.title.y = element_text(size=rel(.8)),plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.5)))

print(p)

```

###Eccentricity Variance


```{r, eval=FALSE}
DT <- DT[, EccentricityVariance := var(Nuclei_CP_AreaShape_Eccentricity_RZSNorm), by="Barcode,ECMpAnnotID"]

p <-  ggplot(DT, aes(x=as.factor(Barcode), y = EccentricityVariance, colour = ECMpAnnotID))+
  geom_point(size = rel(1.5), alpha = .8)+
  ggtitle("Eccentricity Variance In Control Wells")+
  xlab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)),axis.title.x = element_text(size=rel(.8)),axis.title.y = element_text(size=rel(.8)),plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.5)))
print(p)

p <-  ggplot(DT, aes(x=as.factor(Barcode), y = log2(EccentricityVariance), colour = SS))+
  geom_violin()+
  ggtitle("Eccentricity Variance In Control Wells")+
  xlab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)),axis.title.x = element_text(size=rel(.8)),axis.title.y = element_text(size=rel(.8)),plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.5)))
print(p)


```

###Area Variance


```{r, eval=FALSE}
DT <- DT[, AreaVariance := var(Nuclei_CP_AreaShape_Area_RZSNorm), by="Barcode,ECMpAnnotID"]

p <-  ggplot(DT, aes(x=as.factor(Barcode), y = AreaVariance, colour = ECMpAnnotID))+
  geom_point(size = rel(1.5), alpha = .8)+
  ggtitle("Area Variance In Control Wells")+
  xlab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)),axis.title.x = element_text(size=rel(.8)),axis.title.y = element_text(size=rel(.8)),plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.5)))
print(p)

p <-  ggplot(DT, aes(x=as.factor(Barcode), y = log2(AreaVariance), colour = SS))+
  geom_violin()+
  ggtitle("Area Variance In Control Wells")+
  xlab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)),axis.title.x = element_text(size=rel(.8)),axis.title.y = element_text(size=rel(.8)),plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.5)))
print(p)

```


###Median DAPI Intensity Variance


```{r, eval=FALSE}
DT <- DT[, MedianDAPIVariance := var(Nuclei_CP_Intensity_MedianIntensity_Dapi_RZSNorm), by="Barcode,ECMpAnnotID"]

p <-  ggplot(DT, aes(x=as.factor(Barcode), y = log2(MedianDAPIVariance), colour = ECMpAnnotID))+
  geom_point(size = rel(1.5), alpha = .8)+
  xlab("")+
  #ylim(0,.01)+
  ggtitle("Median DAPI Intensity Variance In Control Wells")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)),axis.title.x = element_text(size=rel(.8)),axis.title.y = element_text(size=rel(.8)),plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.5)))
print(p)


```


```{r, eval=FALSE}

p <-  ggplot(DT, aes(x=as.factor(Barcode), y = log2(MedianDAPIVariance), colour = SS))+
  geom_violin()+
  xlab("")+
    #ylim(0,.005)+
  ggtitle("Median DAPI Intensity Variance In Control Wells")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)),axis.title.x = element_text(size=rel(.8)),axis.title.y = element_text(size=rel(.8)),plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.5)))
print(p)

```

##Selected Well  Boxplots by ECM Protein
The following boxplots show the rank-ordered individual ECM protein responses of selected feature vectors stratified by ECM protein in a selected well.

Eccentricity: The eccentricity of the ellipse that has the same second-moments as the region. The eccentricity is the ratio of the distance between the foci of the ellipse and its major axis length. The value is between 0 and 1. (0 and 1 are degenerate cases; an ellipse whose eccentricity is 0 is actually a circle, while an ellipse whose eccentricity is 1 is a line segment.)   

Extent: The proportion of the pixels in the bounding box that are also in the region. Computed as the Area divided by the area of the bounding box.  

```{r, eval=FALSE}
p <-  ggplot(DT[DT$Barcode == "LI8X00402"], aes(x=reorder(ECMpAnnotID, Nuclei_CP_AreaShape_Eccentricity_RZSNorm, FUN=median), y = Nuclei_CP_AreaShape_Eccentricity_RZSNorm, colour = SS))+
  geom_boxplot()+
  #xlim(1,20)+ylim(35,1)
  #scale_color_gradient(low="white",high="red")+
  facet_wrap(~Barcode, ncol=4)+
  ggtitle("Eccentricity In Control Wells")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)),axis.title.x = element_text(size=rel(.8)),axis.title.y = element_text(size=rel(.8)),plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.5)))

print(p)
```

```{r, eval=FALSE}

p <-  ggplot(DT[DT$Barcode == "LI8X00402"], aes(x=reorder(ECMpAnnotID, Nuclei_CP_AreaShape_Extent_RZSNorm, FUN=median), y = Nuclei_CP_AreaShape_Extent_RZSNorm, colour = SS))+
  geom_boxplot()+
  #xlim(1,20)+ylim(35,1)
  #scale_color_gradient(low="white",high="red")+
  facet_wrap(~Barcode, ncol=4)+
  ggtitle("Extent In Control Wells")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)),axis.title.x = element_text(size=rel(.8)),axis.title.y = element_text(size=rel(.8)),plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.5)))

print(p)

```

##Cell Count
The following plots show spot cell count vs the array rows and then the array columns. These show the location and magnitude of the differences in cell counts.  


```{r, eval=FALSE, fig.height=20, fig.width=12, , eval=FALSE}

p <-  ggplot(DT, aes(x=Spot_PA_SpotCellCount, y = ArrayRow))+
  geom_point(size = .8)+
  ggtitle(titleName)+
  facet_wrap(~Barcode, ncol=4)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)),axis.title.x = element_text(size=rel(.8)),axis.title.y = element_text(size=rel(.8)),plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.5)))
print(p)


```


```{r, eval=FALSE, fig.height=20, fig.width=12, , eval=FALSE}

p <-  ggplot(DT, aes(x = ArrayColumn, y=Spot_PA_SpotCellCount))+
  geom_point(size = .8)+
  ggtitle(titleName)+
  facet_wrap(~Barcode, ncol=4)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)),axis.title.x = element_text(size=rel(.8)),axis.title.y = element_text(size=rel(.8)),plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.5)))
print(p)


```

[comment]: <> (##Print Block Effects (in development)
The following boxplots show the spot cell counts for each printblock in each well.
)

```{r, fig.height=20, fig.width=12, eval=FALSE}
#DEBUG: need to bring block through as a column that doesn't get a CV

p <-  ggplot(CVDT[CVDT$Order=="Actual"], aes(x=as.factor(Barcode), y = Spot_PA_SpotCellCount))+
  geom_boxplot()+
  #xlim(1,20)+ylim(35,1)+
  #scale_color_gradient(low="white",high="red")+
  facet_wrap(~Block, ncol=7)+
  ggtitle("CVs of Spot Cell Count In Control Wells")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)),axis.title.x = element_text(size=rel(.8)),axis.title.y = element_text(size=rel(.8)),plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.5)))

print(p)


```

```{r zPrimeOfWells, eval=FALSE}

ZPFList <- lapply(unique(CVDT$Barcode), function(barcode, features, dt){
  #browser()
  setkey(dt,Barcode)
  bDT <-dt[barcode]
  zpList <- lapply(features, function(feature, bDT){
    #browser()
    pos <- bDT[[feature]][bDT$Order=="Actual" ]
    neg <- bDT[[feature]][bDT$Order=="Random" ][sample(1:sum(bDT$Order=="Random"),size = length(pos), replace=FALSE)]
    zp <- data.table(Feature = feature, 
                     ZPrimeFactor = calculateZPrime(pos, neg),
                     KSPvalue = ks.test(x = pos, y = neg)[["p.value"]],
                     bh.distance = separability.measures(Vector.1 = pos, Vector.2 = neg)[["bh"]])
    return(zp)
  },bDT = bDT)
  zpDT <- rbindlist(zpList)
  zpDT$Barcode <- barcode
  return(zpDT)
},dt = CVDT,
features =grep("Eccentricity|Extent|MedianIntensity_Dapi|Area$|SpotCellCount$",colnames(CVDT), value = TRUE))

ZPFDT <- rbindlist(ZPFList)


p <- ggplot(ZPFDT, aes(x = as.factor(Barcode), y = ZPrimeFactor, colour = Feature))+
  geom_point(size = 5, alpha = .8)+
  ggtitle("Z Prime Factors of CVs In Control Wells")+
  ylim(-200,1)+
  xlab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(1)),axis.title.x = element_text(size=rel(1)),axis.title.y = element_text(size=rel(.8)),plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.8)),legend.title=element_text(size = rel(1)))
print(p)


p <- ggplot(ZPFDT, aes(x = as.factor(Barcode), y = KSPvalue, colour = Feature))+
  geom_point(size = 5, alpha = .8)+
  ggtitle("KS P-values of CVs In Control Wells")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)),axis.title.x = element_text(size=rel(.8)),axis.title.y = element_text(size=rel(.8)),plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.8)),legend.title=element_text(size = rel(.8)))
print(p)

p <- ggplot(ZPFDT, aes(x = as.factor(Barcode), y = bh.distance, colour = Feature))+
  geom_point(size = 5, alpha = .8)+
  ggtitle("BH Distance of CVs In Control Wells")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)),axis.title.x = element_text(size=rel(.8)),axis.title.y = element_text(size=rel(.8)),plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.8)),legend.title=element_text(size = rel(.8)))
print(p)

```

##Supplemental Material

###Rank Ordered Eccentricity

This table lists the ECM proteins in the 402 A03 well, ordered by summarized eccentricity response. This table identifies ECM proteins that cause different eccentricity responses. It is followed by a table that ranks each spot by eccentricity and gives the coordinates of the images.  

```{r, eval=FALSE}
setkey(DT4,Nuclei_CP_AreaShape_Eccentricity_RZSNorm)
kable(DT4[DT4$Barcode=="LI8X00402", list(ECMpAnnotID,Nuclei_CP_AreaShape_Eccentricity_RZSNorm)], digits = 2)
```

```{r, eval=FALSE}

setkey(DT,Nuclei_CP_AreaShape_Eccentricity_RZSNorm)
kable(DT[DT$Barcode=="LI8X00402", list(ECMpAnnotID,ArrayRow,ArrayColumn,Nuclei_CP_AreaShape_Eccentricity_RZSNorm)], digits = 2)
```


```{r zPrimeToCOLI, eval=FALSE}

eccentricityZPFList <- lapply(unique(DT$ECMpAnnotID[DT$Barcode=="LI8X00402"]), function(ECMp, DT){
  pos <- DT$Nuclei_CP_AreaShape_Eccentricity[DT$ECMpAnnotID==ECMp ]
  neg <- DT$Nuclei_CP_AreaShape_Eccentricity[DT$ECMpAnnotID=="COL1_Own" ]
  zp <- calculateZPrime(pos, neg)
  names(zp) <- ECMp
  return(zp)
},DT = DT[DT$Barcode=="LI8X00402"])

zpDT <- cbind(names(unlist(eccentricityZPFList)),unlist(eccentricityZPFList))

p <- ggplot(DT[DT$Barcode=="LI8X00402"], aes(x = Nuclei_CP_AreaShape_Eccentricity))+
  geom_density(fill = "blue",binwidth = .02, alpha = .5)+
  facet_wrap(~ECMpAnnotID, ncol=4)

p <-  p+ geom_density(data = DT[DT$Barcode=="LI8X00402" & DT$ECMpAnnotID == "COL1_Own"], aes(x = Nuclei_CP_AreaShape_Eccentricity), fill = "black", binwidth = .02, alpha = .5)
print(p)

```

```{r, eval=FALSE}
ss <- "SS1"
l1 <- fread(paste0("./",cellLine,"/",ss,"/AnnotatedData/",cellLine,"_",ss,"_Level1.txt"), showProgress = FALSE)
setkey(l1, Ligand)
l1 <- l1["FBS"]
    setkey(l1,"ECMp")
    l1 <- l1[!"fiducial"]
    l1 <- l1[!"blank"]

p <- ggplot(l1, aes(x = log2(Cytoplasm_CP_Intensity_MedianIntensity_Actin), y = log2(Nuclei_CP_Intensity_MedianIntensity_Dapi)))+
  geom_point(size = rel(.4))+
  facet_wrap(~Barcode, ncol = 4)
print(p)

ss <- "SS2"
l1 <- fread(paste0("./",cellLine,"/",ss,"/AnnotatedData/",cellLine,"_",ss,"_Level1.txt"), showProgress = FALSE)
setkey(l1, Ligand)
l1 <- l1["FBS"]
    setkey(l1,"ECMp")
    l1 <- l1[!"fiducial"]
    l1 <- l1[!"blank"]

p <- ggplot(l1, aes(x = log2(Nuclei_CP_Intensity_MedianIntensity_H3), y = log2(Nuclei_CP_Intensity_MedianIntensity_Dapi)))+
  geom_point(size = rel(.4))+
  facet_wrap(~Barcode, ncol = 4)
print(p)

ss <- "SS3"
l1 <- fread(paste0("./",cellLine,"/",ss,"/AnnotatedData/",cellLine,"_",ss,"_Level1.txt"), showProgress = FALSE)
setkey(l1, Ligand)
l1 <- l1["FBS"]
    setkey(l1,"ECMp")
    l1 <- l1[!"fiducial"]
    l1 <- l1[!"blank"]

p <- ggplot(l1, aes(x = log2(Cytoplasm_CP_Intensity_MedianIntensity_KRT5), y = log2(Nuclei_CP_Intensity_MedianIntensity_Dapi)))+
  geom_point(size = rel(.4))+
  facet_wrap(~Barcode, ncol = 4)
print(p)


```