
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,
                      echo=FALSE, warning=FALSE, message=FALSE)
```


```{r }
#Author: Mark Dane, copyright 2015

source("./MEPLINCSFunctions.R")

signalDistributions <- function(dt, signal){
  barcodes <- unique(dt$Barcode)
  for (barcode in barcodes){
    DT <- dt[dt$Barcode == barcode] 
    xlimits <- quantile(unlist(DT[,signal, with=FALSE]), probs=c(0,.99))
    
    p <- ggplot(DT, aes_string(x=signal))+
      geom_bar(binwidth=.0005)+
      coord_cartesian(xlim=xlimits)+
      facet_wrap(~Well_Ligand, nrow=2, scales="fixed")+
      ggtitle(paste(cellLine, barcode))
    print(p)  
  }
}

spectralPlots <- function(dt, xSignal, ySignal){
  barcodes <- unique(dt$Barcode)
  for (barcode in barcodes){
    DT <- dt[dt$Barcode == barcode] 
    xlimits <- quantile(unlist(DT[,xSignal, with=FALSE]), probs=c(0,.98))
    ylimits <- quantile(unlist(DT[,ySignal, with=FALSE]), probs=c(0,.98))
    
    p <- ggplot(DT, aes_string(x=xSignal, y=ySignal))+
      geom_point(size=rel(.5), alpha=.5)+
      geom_rug(col=rgb(.5,0,0,alpha=.01))+
      coord_cartesian(xlim=xlimits,ylim=ylimits)+
      facet_wrap(~Well_Ligand, nrow=2, scales="fixed")+
      ggtitle(paste(cellLine, barcode))
    print(p)  
  }
}

signalvsSCC <- function(dt, ySignal){
  barcodes <- unique(dt$Barcode)
  xSignal <- "Spot_PA_SpotCellCount"
  
  for (barcode in barcodes){
    DT <- dt[dt$Barcode == barcode] 
    DT <- DT[DT$Nuclei_CP_Intensity_MedianIntensity_Dapi > .02]
    xlimits <- quantile(unlist(DT[,xSignal, with=FALSE]), probs=c(0,.98))
    ylimits <- quantile(unlist(DT[,ySignal, with=FALSE]), probs=c(0,.98))
    
    p <- ggplot(DT, aes_string(x=xSignal, y=ySignal))+
      geom_point(size=rel(.5), alpha=.5)+
      geom_rug(col=rgb(.5,0,0,alpha=.01))+
      stat_smooth(method = "loess")+
      coord_cartesian(xlim=xlimits,ylim=ylimits)+
      facet_wrap(~Well_Ligand, nrow=2, scales="fixed")+
      ggtitle(paste(cellLine, barcode))
    print(p)  
  }
}
```



```{r setup}
library("ggplot2")
library("data.table")
library("MEMA")
library("grid")
library("knitr")
library("gplots")
library("RColorBrewer")
library(DT)
library(d3heatmap)
library(plotly)

#Setup colors for Barcode and text in all heatmaps
selDark2 <- colorRampPalette(brewer.pal(8,"Dark2"))
plateCol = selDark2(8)
hmcols<-colorRampPalette(c("blue","white","red"))(256)
# analysisVersion <- "v1"
# cellLine <- "PC3"
# ss <- "SS3"

l1 <- fread(paste0("./",cellLine,"/",ss,"/AnnotatedData/",cellLine,"_",ss,"_",analysisVersion,"_Level1.txt"), showProgress = FALSE)

l3 <- fread(paste0("./",cellLine,"/",ss,"/AnnotatedData/",cellLine,"_",ss,"_",analysisVersion,"_Level3.txt"), showProgress = FALSE)

barcodes <- sort(unique(l1$Barcode))


#Set a threshold for filtering wells on their QA score
wellQAThresh <- 0.7

#TODO: Read this from Level 3 data
lthresh <- 0.6

#Number of PCS components to use
nrPCs <- 9

#Z score threshold for extreme spot cell count
SCCZscoreThresh <- 3

#Spot cell count threshold for HF dataset
HFSCCThresh <- 20

#Replicate count threshold for HF Dataset
HFRepThresh <- 3

```
#MEP-LINCS `r cellLine` `r ss` Spectral Analysis
####date: `r Sys.Date()`

##Introduction

This report focuses on the DAPI, `r unique(l3$Endpoint488)` (488nm), `r unique(l3$Endpoint555)` (555nm) and `r unique(l3$Endpoint647)` (647nm) intensity signals in staining set `r ss` in the version `r analysisVersion` dataset of the `r cellLine` cell line. The report starts with histograms of the cell-level signals in each well. These are followed by cell-level biplots of each signal paired with its neighboring channel and then spot level plots of each signal vs the spot cell count.  


###Cell Cycle Plots
Cell Cycle plots include univariate plots of the total DAPI signal. The blue vertical line shows the auto-gate for classifying 2N and 4N populations.


```{r Cell_cycle_plots,echo=FALSE, fig.width=8, fig.height=5, eval=TRUE}

plotTotalDAPI(l1, barcodes)

if (ss == "SS1"){
  
  
} else if (ss == "SS2"){
  for (barcode in barcodes){
    mDT <- l1[l1$Barcode == barcode] 
    mDT <- mDT[mDT$Nuclei_CP_Intensity_IntegratedIntensity_Dapi > quantile(mDT$Nuclei_CP_Intensity_IntegratedIntensity_Dapi, probs=.01, na.rm=TRUE) & mDT$Nuclei_CP_Intensity_IntegratedIntensity_Dapi < quantile(mDT$Nuclei_CP_Intensity_IntegratedIntensity_Dapi,probs=.98, na.rm=TRUE)]
    
    p <- ggplot(mDT, aes(x=Nuclei_CP_Intensity_IntegratedIntensity_Dapi, y = log2(Nuclei_CP_Intensity_MedianIntensity_Edu), colour = Nuclei_PA_Gated_EduPositive))+geom_point(size = rel(.6), alpha = .8)+facet_wrap(~Well_Ligand, nrow=2, scales="free_x")+ggtitle(paste("\n\n","Edu vs. Total DAPI Signal,",barcode))+ylab("Edu (log2)")+xlab("Total Intensity DAPI")+
      theme(strip.text = element_text(size = 5))
    
    suppressWarnings(print(p))
    
  }
} 
```

The rest of the histograms are of the cell-level median intensity of each signal.  


```{r SS3Distributions, eval=ss=="SS3"}
signalDistributions(l1,"Cytoplasm_CP_Intensity_MedianIntensity_KRT5")
signalDistributions(l1,"Cytoplasm_CP_Intensity_MedianIntensity_KRT19")
signalDistributions(l1,"Cytoplasm_CP_Intensity_MedianIntensity_CellMask")

```

##Spectral Overlap
The following plots are intended to show if there is overlap between the channels. Each dot is a cell. The axes are the median intensity of two signals that border in emission frequency.  

```{r SpectralOverlap}
p <- spectralPlots(dt = l1, xSignal = "Nuclei_CP_Intensity_MedianIntensity_Dapi", ySignal = "Cytoplasm_CP_Intensity_MedianIntensity_KRT5")

p <- spectralPlots(dt = l1, xSignal = "Cytoplasm_CP_Intensity_MedianIntensity_KRT5", ySignal = "Cytoplasm_CP_Intensity_MedianIntensity_KRT19")

p <- spectralPlots(dt = l1, xSignal = "Cytoplasm_CP_Intensity_MedianIntensity_KRT19", ySignal = "Cytoplasm_CP_Intensity_MedianIntensity_CellMask")
```

###Signals vs Spot Cell Count
The median intensity of each signal at the cell level is median summarized to the spot level then plotted vs the spot cell count.  
```{r DAPIvsSCC}
signalvsSCC(l3, ySignal = "Nuclei_CP_Intensity_MedianIntensity_Dapi")
signalvsSCC(l3, ySignal = "Cytoplasm_CP_Intensity_MedianIntensity_KRT5")
signalvsSCC(l3, ySignal = "Cytoplasm_CP_Intensity_MedianIntensity_KRT19")
signalvsSCC(l3, ySignal = "Cytoplasm_CP_Intensity_MedianIntensity_CellMask")

```